### 2021-12-27 paper/markdown/figures.Rmd
Add files via upload

---
title: "Radio CovILD, 1-year follow-up"
subtitle: "Tables and figures"
author: "CovILD study team"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
always_allow_html: true
output: 
  pdf_document:
    latex_engine: xelatex
bibliography: cov_biblio.bib
csl: elife.csl
header-includes:
  \usepackage{longtable}
  \usepackage{tabu}
  \usepackage{caption}
  \usepackage{makecell}
  \usepackage{pdflscape}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage{wrapfig}
  \usepackage{multirow}
  \usepackage[normalem]{ulem}
  \usepackage{colortbl}
  \usepackage{xcolor}
  \usepackage{float} \floatplacement{figure}{H}
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}       \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---

```{r, setup, include = FALSE}

library(bookdown)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, dev = "cairo_pdf", ft_max_row = 6, tab.topcaption = FALSE, ft.align = "center")

set_flextable_defaults(font.size = 11, padding = 2)

use_df_printer()

```

\newpage

# Tables

```{r tab-study-variables, tab.cap = 'Variables available for the analysis. The first 10 table records are shown. The full table is available as a supplementary Excel file.'}

paper_tables$study_vars[1:10, ] %>% 
  flextable() %>% 
  width(1, width = 4, unit = 'cm') %>% 
  width(2:3, width = 2.5, unit = 'cm') %>% 
  width(4:5, width = 4, unit = 'cm')

```

\newpage

```{r tab-baseline-cohort, tab.cap = 'Baseline characteristic of the cohort.'}

paper_tables$baseline_cohort %>% 
  flextable() %>%
  width(1, width = 3, unit = 'cm') %>% 
  width(2, width = 7, unit = 'cm')

```

\newpage

```{r tab-baseline-severity, tab.cap = 'Baseline characteristic of the acute COVID-19 severity subsets. Statistical significance of the differences between the severity subsets was assessed with $\\chi^2$ test.'}

paper_tables$baseline_severity %>% 
  flextable() %>%
  width(1, width = 1.8, unit = 'cm') %>% 
  width(2:5, width = 3.9, unit = 'cm') %>% 
  width(6, width = 2.5, unit = 'cm')

```

\newpage

```{r tab-ct-cohort, tab.cap = 'Lung CT features in the entire cohort at the consecutive follow-up visits. Statistical significance of the differences between the visits was assessed with $\\chi^2$ test.'}

paper_tables$ct_cohort %>% 
  flextable() %>%
  width(1, width = 3, unit = 'cm') %>% 
  width(2:5, width = 3.2, unit = 'cm') %>% 
  width(6, width = 2.5, unit = 'cm')

```

\newpage

```{r tab-ct-visits, tab.cap = 'Lung CT features in the acute COVID-19 severity subsets at the consecutive follow-up visits. Statistical significance of the differences between the visits was assessed with $\\chi^2$ test.'}

paper_tables$ct_visits %>% 
  flextable() %>%
  width(1, width = 1.8, unit = 'cm') %>% 
  width(2, width = 2.7, unit = 'cm') %>% 
  width(3:6, width = 3.2, unit = 'cm') %>% 
  width(7, width = 2.5, unit = 'cm')

```

\newpage

```{r tab-ct-severity, tab.cap = 'Lung CT features in the acute COVID-19 severity subsets at the consecutive follow-up visits. Statistical significance of the differences between the severity subsets was assessed with $\\chi^2$ test.'}

paper_tables$ct_severity %>% 
  flextable() %>%
  width(1, width = 1.8, unit = 'cm') %>% 
  width(2, width = 2.7, unit = 'cm') %>% 
  width(3:6, width = 3.2, unit = 'cm') %>% 
  width(7, width = 2.5, unit = 'cm')

```

\newpage

```{r tab-ctss, tab.cap = 'CT severity score at the consecutive visits in the entire cohort and the acute COVID-19 severity subsets. Statistical significance of the differences between the visits was assessed with Kruskal-Wallis test.'}

paper_tables$ctss %>% 
  flextable() %>%
  width(1, width = 1.8, unit = 'cm') %>% 
  width(2:5, width = 3.5, unit = 'cm') %>% 
  width(6, width = 2.5, unit = 'cm')

```

\newpage

# Figures {#sec:figures}

```{r fig-consort, fig.width = mm_inch(paper_figures$ct_kinetic$w), fig.height = mm_inch(paper_figures$ct_kinetic$h), dpi = 600, fig.cap = 'Study inclusion flow diagram.'}

ggdraw()

```

__Figure \@ref(fig:fig-consort). Study inclusion flow diagram.__

\newpage

```{r fig-ct-kinetic, fig.width = mm_inch(paper_figures$ct_kinetic$w), fig.height = mm_inch(paper_figures$ct_kinetic$h), dpi = 600, fig.cap = 'Resolution kinetic of any and moderate-to-severe lung CT abnormalities.'}

paper_figures$ct_kinetic$plot

```

__Figure \@ref(fig:fig-ct-kinetic). Resolution kinetic of any and moderate-to-severe lung CT abnormalities.__

Risk of any lung CT abnormalities (__A__) and moderate-to-severe lung CT abnormalities (CTSS > 5, __B__) at the consecutive time points was modeled by mixed-effect logistic regression (random effect: individual, fixed effect: time post COVID-19) in the entire cohort and the acute COVID-19 severity subsets. Significance of the fixed time term was determined by likelihood ratio test (LRT) versus the respective random effect-only null model. Plots display percents of individuals with CT abnormalities in the analyzed collectives. $\chi$ statistic, degrees of freedom and p values of the LRT test are presented in the plot captions. Numbers of individuals with the complete set of consecutive CT scans are shown under the plots.

\newpage

```{r fig-ctss-kinetic, fig.width = mm_inch(paper_figures$ctss_kinetic$w), fig.height = mm_inch(paper_figures$ctss_kinetic$h), dpi = 600, fig.cap = 'Kinetic of the CT severity score.'}

paper_figures$ctss_kinetic$plot

```

__Figure \@ref(fig:fig-ctss-kinetic). Kinetic of the CT severity score.__

CTSS kinetic at the consecutive time points was investigated by Friedman test (grouping by the individual) in the entire cohort and the acute COVID-19 severity subsets. The effect size was determined by Kendall W test, differences between particular time points were compared by Wilcoxon test. Plots display individual CTSS value trajectories as thin gray lines, thick colored lines represent medians, interquartile ranges are presented as colored regions. $\chi$ statistic, degrees of freedom and p values of the Friedman test and the Kendall W statistic with 95$%$ confidence intervals are presented in the plot captions. Numbers of individuals with the complete set of consecutive CT scans are shown under the plots.

\newpage

```{r fig-pili-risk, fig.width = mm_inch(paper_figures$pili_modeling$w), fig.height = mm_inch(paper_figures$pili_modeling$h), dpi = 600, fig.cap = 'Risk of developing PILI at the one-year post-COVID-19 follow-up.'}

paper_figures$pili_modeling$plot

```

__Figure \@ref(fig:fig-pili-risk). Risk of developing PILI at the one-year post-COVID-19 follow-up.__

__(A)__ Risk factors of developing PILI at the one-year follow-up visit were identified by a series of logistic models. Odds ratio (OR) significance was determined by Wald Z test. OR with 95$%$ confidence intervals are presented in a Forest plot. Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.

__(B)__ Independent risk factors of PILI were identified by multi-parameter logistic modeling with backward elimination. OR significance was determined by Wald Z test. Model prediction was verified by 20-fold cross-validation (CV) and receiver-operating characteristic (ROC). OR with 95$%$ confidence intervals are presented in a Forest plot. Area under the curve (AUC), sensitivity (Se) and specificity (Sp) for detection of PILI in CV are indicated in the plot caption. Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.

BMI: body mass index, PKY: packages per year.

\newpage

```{r fig-ctss-model, fig.width = mm_inch(paper_figures$ctss_modeling$w), fig.height = mm_inch(paper_figures$ctss_modeling$h), dpi = 600, fig.cap = 'Predictors of CTSS at the one-year post-COVID-19 follow-up.'}

paper_figures$ctss_modeling$plot

```

__Figure \@ref(fig:fig-ctss-model). Predictors of CTSS at the one-year post-COVID-19 follow-up.__

__(A)__ Factors associated with CTSS at the one-year follow-up visit were identified by a series of Poisson models. Regression coefficient ($\beta$) significance was determined by Wald Z test. Exponent $\beta$ with 95$%$ confidence intervals are presented in a Forest plot. Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.

__(B)__ Independent factors associated with CTSS were identified by multi-parameter Poisson modeling with backward elimination. $\beta$ significance was determined by Wald Z test. Model prediction was verified by 20-fold cross-validation (CV). Exponent $\beta$ with 95$%$ confidence intervals are presented in a Forest plot. Mean absolute error (MAE) for prediction of CTSS and model R^2^ (Rsq) in CV are indicated in the plot caption. Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.

BMI: body mass index, PKY: packages per year.

### 2021-12-27 analysis scripts/eda.R

@@ -1,0 +1,193 @@
+# This script performs explorative data analysis: baseline characteristic
+# and frequency of CT findings at the particular visits. Finally, distribution
+# statistics of the CT severity score are calculated
+
+  insert_head()
+  
+# container list ------
+  
+  eda <- list()
+  
+# baseline clinical characteristic -----
+  
+  insert_msg('Baseline clinics, entire cohort')
+  
+  ## analyses
+
+  eda$baseline$analyses_cohort <- radio$long_expl_var %>% 
+    map(analyze_feature, 
+        inp_tbl = rbind(mutate(radio$clear, dummy_var = factor('dummy1')), 
+                        mutate(radio$clear, dummy_var = factor('dummy2'))), 
+        split_var = 'dummy_var') %>% 
+    set_names(radio$long_expl_var)
+
+  ## summary table
+  
+  eda$baseline$summary_cohort <- eda$baseline$analyses_cohort %>% 
+    map_dfr(get_feature_summary) %>% 
+    mutate(label = translate_var(variable), 
+           statistic = dummy1) %>% 
+    select(variable, label, statistic)
+  
+# baseline characteristic in the severity groups -----
+  
+  insert_msg('Baseline clinics, severity groups')
+  
+  ## analyses
+  
+  eda$baseline$analyses_severity <- radio$long_expl_var %>% 
+    map(analyze_feature, 
+        inp_tbl = radio$clear, 
+        split_var = 'severity') %>% 
+    set_names(radio$long_expl_var)
+  
+  ## summary tables, for age: ANOVA (homogeneous variances and normality, see the analysis object)
+  
+  eda$baseline$summary_severity <- eda$baseline$analyses_severity %>% 
+    map_dfr(get_feature_summary) %>% 
+    filter(variable != 'severity') %>% 
+    mutate(p_value = ifelse(is.na(p_chi), p_param, p_chi)) %>% 
+    add_stat_info %>% 
+    select(variable, label, mild, moderate, severe, critical, test, p_value, significance, p_adj)
+  
+# CT features at the time points, entire cohort ------
+  
+  insert_msg('CT features, entire cohort')
+  
+  eda$ct$analyses_cohort <- radio$cat_outcome_var %>% 
+    map(~analyze_feature(inp_tbl = radio$long[c('visit', .x)] %>% 
[…]

### 2021-12-27 analysis scripts/kinetic.R

@@ -1,0 +1,146 @@
+# This script 'models' kinetic of the CT score in the entire cohort and severity groups
+# by Friedman test.
+# Risk of incomplete resolution of any, moderate-to-severe, severe and particular type abnormalities
+# is estimated by mixed-effect logistic modeling. Only the participants with the complete longitudinal CT scan sets are analyzed
+
+  insert_head()
+  
+# container list -----
+  
+  kinetic <- list()
+  
+# analysis table ----
+  
+  kinetic$analysis_tbl <- radio$long %>% 
+    filter(ID %in% radio$complete_long_ids)
+  
+  kinetic$analysis_tbl <- kinetic$analysis_tbl %>% 
+    dlply(.(severity)) %>% 
+    c(list(cohort = kinetic$analysis_tbl), .)
+  
+# CTSS kinetics -----
+  
+  insert_msg('CTSS kinetic')
+  
+  ## test
+  
+  kinetic$ctss$analysis <- kinetic$analysis_tbl %>% 
+    map(~friedman_test(ctss ~ visit|ID, data = .x))
+  
+  ## summary
+  
+  kinetic$ctss$summary <- kinetic$ctss$analysis %>% 
+    map2_dfr(., names(.), ~mutate(.x, subset = .y, p_value = p)) %>% 
+    mutate(p_adj = p.adjust(p_value, 'BH'), 
+           plot_caption = paste0('\u03C7 = ', signif(statistic, 2), 
+                                 ', df = 3, p = ', signif(p_value, 2)))
+  
+  ## effect size determined by Kendall W, updating the plot_caption
+  
+  plan('multisession')
+  
+  kinetic$ctss$effsize <- kinetic$analysis_tbl %>% 
+    future_map(~friedman_effsize(ctss ~ visit|ID, 
+                                 data = .x, 
+                                 ci = TRUE, 
+                                 ci.type = 'bca'), 
+               .options = furrr_options(seed = TRUE)) %>% 
+    map2_dfr(., names(.), ~mutate(.x, subset = .y))
+  
+  kinetic$ctss$summary <- kinetic$ctss$summary %>% 
+    mutate(plot_caption = list(x = plot_caption, 
+                               y = kinetic$ctss$effsize$effsize, 
+                               lo = kinetic$ctss$effsize$conf.low, 
+                               hi = kinetic$ctss$effsize$conf.high) %>% 
[…]

### 2021-12-27 analysis scripts/risk_score.R

@@ -1,0 +1,185 @@
+# This script models risk of any, moderate-severe, severe, ggo, reticulation, consolidations and bronchial dilatation 
+# (logistic regression) as well as of the abnormality grading at the 1-year follow-up. Explanatory variables
+# are> sex, age, severity, bin, smoking parameters as well as initial CT abnormality severity (CTSS with the cutoffs 5 and 10)
+# Multi-paramater models are constructed by backward elimination and validated by 20-fold CV
+
+  insert_head()
+  
+# container list -----
+  
+  risk <- list()
+
+
+# globals ------
+  
+  insert_msg('Globals')
+  
+  risk$explanatory_vars <- c(radio$long_expl_var)
+  
+  risk$outcomes <-  c('ggo_fup4', 
+                      'retic_fup4', 
+                      'pili_fup4', 
+                      'ctss_any_fup4', 
+                      'ctss_fup4')
+  
+  risk$outcome_forms <- risk$outcomes %>% 
+    map(~paste(.x, paste(risk$explanatory_vars, collapse = '+'), sep = '~')) %>% 
+    map(as.formula)
+  
+  risk$families <- c(rep('binomial', 4), 'poisson')
+  
+  risk$analysis_tbl <- radio$clear
+  
+  for(i in risk$outcomes) {
+    
+    if(is.factor(risk$analysis_tbl[[i]])) {
+      
+      risk$analysis_tbl <- risk$analysis_tbl %>% 
+        mutate(!!sym(i) := as.numeric(.data[[i]]) - 1)
+      
+    }
+    
+  }
+  
+  risk$levene_forms <- risk$explanatory_vars %>% 
+    paste('.resid ~', .) %>% 
+    map(as.formula)
+  
+# serial univariate modeling -----
+
+  insert_msg('Serial univariate modeling')  
+  
+  risk$univariate_models <- list(response = risk$outcomes, 
+                                 family = risk$families) %>% 
+    pmap(make_lm_model, 
+        data = risk$analysis_tbl, 
+        indep_variable = risk$explanatory_vars, 
+        mod_fun = glm, 
+        est_transf = exp, 
+        error_resistant = FALSE) %>% 
+    set_names(risk$outcomes)
+
+  ## model diagnostic tables, normality and variance homogeneity testing
+  
+  risk$univariate_resids <- risk$univariate_models %>% 
+    map(~map(.x, get_qc_tbl))
+  
[…]

### 2021-12-27 data_analysis.R

@@ -1,0 +1,30 @@
+# This script performs data analysis 
+
+  insert_head()
+  
+# tools -----
+  
+  c('./tools/sys_tools.R', 
+    './tools/project_tools.R', 
+    './tools/project_globals.R', 
+    './tools/counting_tools.R', 
+    './tools/modeling_tools.R', 
+    './tools/lm_qc_tools.R') %>% 
+    walk(source)
+  
+  library(DescTools)
+  library(rstatix)
+  library(ggpubr)
+  library(furrr)
+  library(caret)
+  
+# analysis scripts -----
+  
+  c('./analysis scripts/eda.R', 
+    './analysis scripts/kinetic.R', 
+    './analysis scripts/risk_score.R') %>% 
+    walk(source)
+  
+# END -----
+  
+  insert_tail()

### 2021-12-27 data_import.R

@@ -1,0 +1,158 @@
+# Data import from a SPSS file and wrangling
+
+# tools -----
+
+  library(plyr)
+  library(tidyverse)
+  library(foreign)
+  library(stringi)
+  library(readxl)
+
+  c('./tools/sys_tools.R', 
+    './tools/project_globals.R', 
+    './tools/project_tools.R') %>% 
+    walk(source)
+  
+  insert_head()
+  
+# container list ----
+  
+  radio <- list()
+
+# Reading the SPSS datafile -----
+  
+  insert_msg('Reading the SPSS data file')
+  
+  radio$raw <- read.spss(file = './data/20211216PILI.sav', 
+                          to.data.frame = TRUE) %>% 
+    as_tibble %>% 
+    mutate(ID = rm_spaces(RAD_ID, 'character'))
+
+  
+# Reading the recoding scheme ----
+  
+  insert_msg('Reading the recoding scheme')
+  
+  radio$vars <- read_excel('./data/coding_scheme.xlsx') %>% 
+    mutate(args = list(args1, args2, args3, args4) %>% 
+             pmap(function(a1, a2, a3, a4) list(if(!is.na(a1)) a1 else NULL, 
+                                                if(!is.na(a2)) a2 else NULL, 
+                                                if(!is.na(a3)) a3 else NULL, 
+                                                if(!is.na(a4)) a4 else NULL)), 
+           args = map(args, compact))
+  
+  radio$vars <- left_join(radio$vars, 
+                          tibble(old_var = names(attr(radio$raw, 'variable.labels')), 
+                                 description = attr(radio$raw, 'variable.labels')), 
+                          by = 'old_var')
+  
+# serial clearing -----
+  
+  insert_msg('Serial cleaning')
+  
+  radio$clear <- radio$vars %>% 
+    filter(recode == 'yes') %>% 
+    select(old_var, new_var, trans_fun, args) %>% 
+    pmap(recode_var, 
+         data = radio$raw) %>% 
+    reduce(left_join, by = 'ID')
+  
+  difftime_years(vector1 = radio$raw$DOE_1, vector2 = radio$raw$DOB)
+  
+  radio$raw %>% 
+    mutate(age = difftime_years(DOE_1, DOB))
+  
+# manual re-coding, cutting down to the patients having completed the last visit ----
+  
[…]

### 2021-12-27 exec_all.R

@@ -1,0 +1,61 @@
+# This program executes the project scripts
+
+# libraries -----
+
+  require(readr)
+
+# listing scripts ----
+
+  executable_scripts <- c('data_import.R', 
+                          'data_analysis.R', 
+                          'render_paper.R') 
+
+# error container ----
+  
+  exec_errors <- list()
+  
+# executing the script list ----
+  
+  for(file in executable_scripts) {
+    
+    diagn_output <- try(source(file, encoding = 'UTF-8'), silent = T) ## UTF-8 encoding cause of greek letters 
+                                                                      ## in plots and tables
+    
+    if(class(diagn_output) == 'try-error') {
+      
+      exec_errors <- c(exec_errors, 'execution failed')
+      
+    } else {
+      
+      exec_errors <- c(exec_errors, 'execution successful')
+      
+    }
+
+  }
+  
+  require(dplyr)
+  
+  exec_errors <- exec_errors %>% 
+    reduce(c) %>% 
+    tibble(execution_result = .)
+  
+  exec_errors <- exec_errors %>% 
+    mutate(file = executable_scripts) %>% 
+    select(file, execution_result)
+  
+  rm(diagn_output, executable_scripts)
+  
+  message(rep('-', 60))
+  print(exec_errors)
+  message(rep('-', 60))
+  
+# saving execution results on the disc ----
+
+  exec_errors %>% 
+    write_tsv('exec_log.log')
+  
+# saving the image -----
+  
+  save.image()
+  
+# END ----

### 2021-12-27 paper scripts/deploy_paper.R

@@ -1,0 +1,65 @@
+# This script renders the legends for tables and figures ----
+
+# data and tools ----
+
+  source('./tools/sys_tools.R')
+
+  library(knitr)
+  library(bookdown)
+  library(kableExtra)
+  library(rmarkdown)
+  library(flextable)
+
+  insert_head()
+  
+# Specific functions -----
+  
+  tab_preview <- function(inp_tbl, head_lines = 'all', signif_digits = 3) {
+    
+    ## makes a nice table head for preview in the figure file
+    
+    out_tbl <- inp_tbl %>% 
+      map_dfc(function(x) if(is.numeric(x)) signif(x, digits = signif_digits) else x)
+    
+    if(head_lines != 'all') {
+      
+      return(out_tbl[1:head_lines, ])
+      
+    } else {
+      
+      return(out_tbl)
+      
+    }
+    
+  }
+  
+  collapse_and <- function(vector, collapse = ', ', last_link = ' and ') {
+    
+    ## collapses a given vector with the last_link string before the last item
+    
+    coll_phrase <- paste(vector[1:length(vector) - 1], collapse = collapse) %>% 
+      paste(., vector[length(vector)], sep = last_link)
+    
+    return(coll_phrase)
+    
+  }
+  
+  mm_inch <- function(input_mm) {
+    
+    return(0.0393700787 * input_mm)
+    
+  }
+
+# Generating figure legend file -----
+  
+  insert_msg('Rendering figure and table legend file')
+  
+  render('./paper/markdown/figures.Rmd', 
+         output_format = word_document2(number_sections = F, theme = 'readable'), 
+         output_dir = './paper/')
+
+
+# END -----
+  
+  insert_tail()
+

### 2021-12-27 paper scripts/paper_figures.R

@@ -1,0 +1,116 @@
+# This script generates paper figures -----
+
+  insert_head()
+
+# container list -----
+
+  paper_figures <- list()
+  
+# Figure 1: Consort plot - placeholder, provided by the Author team ----
+  
+  insert_msg('Figure 1: CONSORT, provided by the authors')
+  
+# Figure 2: kinetics of any CT abnormalities -----
+  
+  insert_msg('Figure 2: kinetic of any CT abnormalities and CTSS')
+  
+  paper_figures$ct_kinetic$upper_panel <- kinetic$binary$plots[c('cohort.ctss_any', 
+                                                                 'mild.ctss_any', 
+                                                                 'moderate.ctss_any', 
+                                                                 'severe.ctss_any', 
+                                                                 'critical.ctss_any')] %>% 
+    map2(., globals$subset_labels, 
+         ~.x + 
+           labs(title = .y) + 
+           scale_y_continuous(limits = c(0, 115), 
+                              breaks = seq(0, 100, by = 25)) + 
+           scale_x_continuous(limits = c(2, 13), 
+                              breaks = c(2, 3, 6, 12))) %>% 
+    plot_grid(plotlist = ., 
+              ncol = 3, 
+              align = 'hv')
+  
+  paper_figures$ct_kinetic$bottom_panel <- kinetic$binary$plots[c('cohort.ctss_mod_severe', 
+                                                                   'mild.ctss_mod_severe', 
+                                                                   'moderate.ctss_mod_severe', 
+                                                                   'severe.ctss_mod_severe', 
+                                                                   'critical.ctss_mod_severe')] %>% 
+    map2(., globals$subset_labels, 
+         ~.x + 
+           labs(title = .y) + 
+           scale_y_continuous(limits = c(0, 115), 
+                              breaks = seq(0, 100, by = 25)) + 
+           scale_x_continuous(limits = c(2, 13), 
+                              breaks = c(2, 3, 6, 12))) %>% 
+    plot_grid(plotlist = ., 
+              ncol = 3, 
+              align = 'hv')
+  
+  paper_figures$ct_kinetic <- plot_grid(paper_figures$ct_kinetic$upper_panel, 
+                                        paper_figures$ct_kinetic$bottom_panel, 
+                                        nrow = 2, 
+                                        labels = LETTERS, 
+                                        label_size = 10) %>% 
+    as_figure_object(figure_label = 'figure_2_kinetic_ct', 
+                     w = 180, 
[…]

### 2021-12-27 paper scripts/paper_tables.R

@@ -1,0 +1,130 @@
+# This script generates paper tables
+
+ insert_head()
+ 
+# container list ----
+ 
+ paper_tables <- list()
+ 
+# Table 1: Study variables -----
+ 
+ insert_msg('Table 1: study variables')
+ 
+ paper_tables$study_vars <- radio$vars %>% 
+   select(old_var, 
+          new_var, 
+          label, 
+          description, 
+          modeling) %>% 
+   mutate(modeling = car::recode(modeling, "'no' = 'not used in modeling'; 'long_outcome' = 'outcome'")) %>% 
+   set_names(c('SPSS variable name', 
+               'R variable name', 
+               'R variable label', 
+               'Description', 
+               'Variable type'))
+ 
+# Table 2: baseline cohort characteristic -----
+ 
+ insert_msg('Table 2: baseline cohort characteristic')
+ 
+ paper_tables$baseline_cohort <- eda$baseline$summary_cohort %>% 
+   select(label, statistic) %>% 
+   set_names(c('Variable', 'Statistic'))
+ 
+# Table 3: baseline characteristic, severity groups -----
+ 
+ insert_msg('Table 3: baseline characteristic, severity groups')
+ 
+ paper_tables$baseline_severity <- eda$baseline$summary_severity %>% 
+   select(label, 
+          mild, 
+          moderate, 
+          severe, 
+          critical, 
+          significance) %>% 
+  set_names(c('Variable', 
+              'Mild', 
+              'Moderate',
+              'Severe', 
+              'Critical', 
+              'Significance'))
+ 
+# Table 4: CT abnormalities at the visits, whole cohort ----
+ 
+ insert_msg('Table 4: CT abnormalities, whole cohort, consecutive visits')
+ 
+ paper_tables$ct_cohort <- eda$ct$summary_cohort %>% 
+   select(label, 
+          starts_with('fup'), 
+          significance) %>% 
+   set_names(c('Variable', 
+               '60-day FUP', 
+               '100-day FUP', 
+               '180-day FUP', 
+               '360-day FUP', 
+               'Significance'))
+ 
+# Table 6: CT abnormalities at the visits, severity ----
+ 
+ insert_msg('Table 5: CT abnormalities, severity groups')
+ 
+ paper_tables$ct_visits <- eda$ct$summary_visit %>% 
+   select(severity, 
+          label, 
+          starts_with('fup'), 
+          significance) %>% 
[…]

### 2021-12-27 render_paper.R

@@ -1,0 +1,24 @@
+# This script renders the paper tables and figures
+
+  insert_head()
+  
+# tools ----
+  
+  c('./tools/sys_tools.R', 
+    './tools/project_tools.R', 
+    './tools/plotting_tools.R') %>% 
+    walk(source)
+  
+  library(writexl)
+  
+# paper scripts -----
+  
+  insert_msg('Sourcing the paper scripts')
+  
+  c('./paper scripts/paper_tables.R', 
+    './paper scripts/paper_figures.R') %>% 
+    walk(source)
+  
+# END -----
+  
+  insert_tail()

### 2021-12-27 tools/counting_tools.R

@@ -1,0 +1,1065 @@
+# This simple script contains some tools for counting variable levels
+# And calculating the characteristic of numeric variables
+
+
+# libraries -----
+
+  require(plyr)
+  require(tidyverse)
+  require(stringi)
+  require(ggsignif)
+  require(ggpubr)
+  
+# functions for factors -------
+
+  count_feature <- function(inp_tbl, var_to_count, remove_na = T, .drop = T) {
+    
+    ## calculates the percentage and number of participants with/without the given feature
+    ## the .drop argument specifies if the empty factor levels will be included in the output
+    
+    if(remove_na) {
+      
+      count_tbl <- inp_tbl %>% 
+        filter(!is.na(.data[[var_to_count]]), 
+               .data[[var_to_count]] != 'no_answer')
+      
+    } else {
+      
+      count_tbl <- inp_tbl
+      
+    }
+    
+    feature_counts <- count_tbl %>% 
+      count(.data[[var_to_count]], 
+            .drop = .drop) %>% 
+      mutate(percent = n/sum(n) * 100, 
+             total_n = sum(n))
+    
+    return(feature_counts)
+    
+  }
+
+  count_feature_lst <- function(inp_tbl, var_to_count_vec, remove_na = T, positive_only = T) {
+    
+    ## wrapper for variable vectors
+    
+    count_tbl <- var_to_count_vec %>% 
+      map(count_feature, 
+          inp_tbl = inp_tbl, 
+          remove_na = remove_na) %>% 
+      map(set_names, 
+          c('feature_strata', 
+            'n', 
+            'percent', 
+            'total_n')) %>% 
+      set_names(var_to_count_vec) %>% 
+      map2_dfr(., names(.), 
+               function(x, y) mutate(x, feature = y))
+    
+    if(positive_only) {
+      
+      count_tbl <- count_tbl %>% 
+        filter(feature_strata == 'yes')
+      
+    }
+    
+    return(count_tbl)
+    
+  }
+  
+  pie_feature <- function(perc_tbl, plot_title = names(perc_tbl)[1], plot_subtitle = NULL, 
+                          order_tbl = NULL, pie = T, repel = T, fill_colors = NULL, cust_theme = NULL) {
+    
+    ## plots the distribution of the given feature: takes an output of the count_feature function
+    ## a tibble with order for plotting may be provided (names: plot_order, feature_strata)
+    
+    plotting_tbl <- perc_tbl %>% 
+      set_names(c('feature_strata', 
[…]

### 2021-12-27 tools/lm_qc_tools.R

@@ -1,0 +1,483 @@
+# This script provides tools for summaries and QC of linear (lm) and generalized linear models
+
+# libraries ------
+
+  require(tidyverse)
+  require(ggrepel)
+  require(broom)
+  require(furrr)
+  require(stringi)
+
+# auxiliary functions ----
+
+  point_plot_ <- function(data, x_var, y_var, x_lab = x_var, y_lab = y_var, 
+                          plot_title = NULL, smooth = T, silent = T, ...) {
+    
+    ## draws a simple point plot for diagnostic purposes, takes the output of get_qc_tbl() as data argument
+    ## color-codes model missfits
+    
+    ## table for plotting 
+    
+    data <- data %>% 
+      mutate(misslab = try(ifelse(.candidate_missfit == 'yes',
+                                  .rownames, 
+                                  NA), silent = T),
+             misslab = ifelse(any(class(misslab) == 'try-error'), NA, misslab))
+    
+    ## fill colors
+    
+    fill_colors <- c(no = 'cornflowerblue', 
+                     yes = 'firebrick4')
+    
+    ## point plot
+    
+    point_plot <- data %>% 
+      ggplot(aes(x = .data[[x_var]], 
+                 y = .data[[y_var]], 
+                 fill = .candidate_missfit)) + 
+      geom_point(size = 2, 
+                 shape = 21) + 
+      geom_text_repel(aes(label = misslab), 
+                      show.legend = F) + 
+      scale_fill_manual(values = fill_colors, 
+                        name = 'Candidate missfit') + 
+      labs(x = x_lab, 
+           y = y_lab, 
+           title = plot_title)
+    
+    if(smooth) {
+      
+      if(silent) {
+        
+        suppressWarnings(point_plot <- point_plot + 
+                           geom_smooth(show.legend = F, 
+                                       color = 'black', 
+                                       fill = 'dodgerblue2', ...))
+        
+      } else {
+        
+        point_plot <- point_plot + 
+          geom_smooth(show.legend = F, 
+                      color = 'black', 
+                      fill = 'dodgerblue2', ...)
+        
+      }
+      
+    }
+      
+    return(point_plot)
+    
+  }
+  
+  calc_expected_ <- function(inp_data, observed) {
+    
+    ## calculates expected normal distribution of a variable observed
[…]

### 2021-12-27 tools/modeling_tools.R

@@ -1,0 +1,526 @@
+# This script provides functional tools for modeling tasks in christas Sepsis project
+
+# libraries -----
+
+  library(plyr)
+  library(dplyr)
+  library(tidyr)
+  library(purrr)
+  library(lme4)
+  library(lmerTest)
+  library(stringi)
+
+# globals ----
+
+  default_signif = 2
+
+# simple model generation functions -----
+
+  make_lmer <- function(inp_table, response = NULL, fix_variables = NULL, random_variables = NULL, interaction_term = '', ...) {
+    
+    # generaes a simple no-interaction lm (if random variables not given) or lmer
+    # Interaction can be included by explicitly stating the interaction term
+    
+    inp_formula <- paste(response, '~', paste(fix_variables, collapse = '+'), interaction_term, sep = '+')
+    
+    if(is.null(random_variables)) {
+      
+      out_model <- lm(formula = as.formula(inp_formula), data = inp_table, ...)
+      
+    } else {
+      
+      for(var in random_variables) {
+        
+        inp_formula <- paste(inp_formula, '+ (1|', var, ')')
+        
+      }
+      
+      out_model <- lmer(formula = as.formula(inp_formula), data = inp_table, ...)
+      
+    }
+    
+    return(out_model)
+    
+  }
+  
+  make_glm <- function(inp_table, response = NULL, fix_variables = NULL, ...) {
+    
+    # generaes a simple no-interaction glm
+    
+    inp_formula <- paste(response, '~', paste(fix_variables, collapse = '+'))
+    
+    return(glm(formula = as.formula(inp_formula), data = inp_table, ...))
+    
+  }
+  
+# model summary functions ----
+
+  get_lme_results <- function(model_mix, L_matrix = NULL, anova = F, ...){
+    
+    require(lmerTest)
+    
+    ## a function returning confidence intervals and significances for particular coefficents of the mixed model with one fixed
+    ## effect. The coefficients to display are defined by the L_matrix. Uses the contest function from the lmerTest package
+    ## If the L_matrix is not given, stats for all coefficients are calculated
+    
+    output <- model_mix %>% 
+      as_lmerModLmerTest
+    
+    if(is.null(L_matrix)) {
+      
+      L_matrix <- identity_matrix(length(model_mix@beta))
+      
+    }
+    
+    if(anova) {
+      
+      output <- contest(output, L_matrix, ...)
+      
+    } else {
+      
+      output <- contest(output, L_matrix, joint = F, ...) %>% 
[…]

### 2021-12-27 tools/project_globals.R

@@ -1,0 +1,56 @@
+# This script sets up graphical and variable globals for the project
+
+  require(plyr)
+  require(tidyverse)
+  
+  insert_head()
+  
+# data container ----
+  
+  globals <- list()
+  
+# Graphics -----
+  
+  insert_msg('Graphics globals')
+  
+  globals$common_text <- element_text(size = 8, face = 'plain', color = 'black')
+  
+  globals$common_margin <- ggplot2::margin(t = 4, l = 3, r = 2, unit = 'mm')
+  
+  globals$common_theme <- theme_classic() + theme(axis.text = globals$common_text, 
+                                                  axis.title = globals$common_text, 
+                                                  plot.title = element_text(size = 8, 
+                                                                            face = 'bold', 
+                                                                            color = 'black', 
+                                                                            hjust = 0), 
+                                                  plot.subtitle = globals$common_text, 
+                                                  plot.tag = element_text(size = 8, 
+                                                                          face = 'plain', 
+                                                                          color = 'black', 
+                                                                          hjust = 0), 
+                                                  plot.tag.position = 'bottom', 
+                                                  legend.text = globals$common_text, 
+                                                  legend.title = globals$common_text, 
+                                                  strip.text = globals$common_text,
+                                                  strip.background = element_rect(fill = 'gray95', color = 'gray80'), 
+                                                  plot.margin = globals$common_margin, 
+                                                  panel.grid.major = element_line(color = 'gray90'))
+  
+# Variable representation -----
+  
+  insert_msg('Variable representation')
+
+  globals$subset_colors <- c('cohort' = 'gray30', 
+                             'mild' = 'steelblue', 
+                             'moderate' = 'cornsilk4', 
+                             'severe' = 'coral3', 
+                             'critical' = 'firebrick')  
+  
+  globals$subset_labels <- c(cohort = 'Cohort', 
+                             mild = 'Mild COVID-19', 
+                             moderate = 'Moderate COVID-19', 
+                             severe = 'Severe COVID-19', 
+                             critical = 'Critical COVID-19')
+  
+# END -----
+  

### 2021-12-27 tools/project_tools.R

@@ -1,0 +1,474 @@
+# This is a container script for the project-specific tools
+
+  require(plyr)
+  require(tidyverse)
+  require(rlang)
+  require(stringi)
+  require(foreign)
+
+  
+# data import and transformation -----
+  
+  recode_var <- function(data, 
+                         old_var, 
+                         new_var, 
+                         ID_var = 'ID', 
+                         trans_fun = 'my_identity', 
+                         args = NULL) {
+    
+    ## renames a variable, if time and id variable names provided, they're included
+    ## in  the output as well
+    ## the transformation function enables
+    ## args specifies the arguments to the transforming function
+    
+    trans_call <- call2(trans_fun, 
+                        data[[old_var]], 
+                        !!!args)
+
+    new_data <- data %>% 
+      mutate(!!ensym(new_var) := eval(trans_call))
+
+    return(new_data[c(ID_var, enexpr(new_var))])
+    
+  }
+  
+  rm_spaces <- function(vector, out_format = 'character') {
+    
+    ## removes trailing spaces
+    
+    cut_vector <- vector %>% 
+      stri_replace_all(regex = '\\s+$', replacement = '')
+    
+    cut_vector <- ifelse(stri_length(cut_vector) == 0, NA, cut_vector)
+    
+    formatter <- switch(out_format,
+                        character = as.character,
+                        numeric = as.numeric,
+                        logical = as.logical)
+    
+    return(formatter(cut_vector, out_format))
+    
+  }
+  
+  recode_vec <- function(vector, recodes, ...) {
+    
+    return(car::recode(vector, 
+                       recodes = recodes, 
+                       ...))
+    
+  }
+  
+  recode_yn <- function(vector, reverse = F, as.factor = T, ...) {
+    
+    ## recodes a 0/1 vector no/yes or the other way round
+    
+    if(reverse) {
+      
+      new_vec <- recode_vec(vector, 
+                            "'yes' = 1;
+                             'no' = 0",
+                            as.factor = as.factor)
+
+    } else {
+      
+      new_vec <- recode_vec(vector, 
+                            "1 = 'yes';
+                             0 = 'no'", 
+                            as.factor = as.factor)
+      
+    }
+
[…]

### 2021-12-27 analysis scripts/eda.R
Add files via upload

@@ -117,8 +117,8 @@
   
   eda$ctss_normality$time_test_ident <- radio$long %>% 
     dlply(.(visit)) %>% 
-    map(~shapiro.test(.x$ctss)) %>% 
-    map2_dfr(., names(.), ~tibble(visit = .y, p_value = .x$p.value))
+    map(~shapiro_test(.x$ctss)) %>% 
+    map2_dfr(., names(.), ~mutate(.x, visit = .y))
   
   eda$ctss_normality$time_plot_ident <- radio$long %>% 
     dlply(.(visit)) %>% 
@@ -130,8 +130,8 @@
   
   eda$ctss_normality$time_sev_test_ident <- radio$long %>% 
     dlply(.(visit, severity)) %>% 
-    map(~shapiro.test(.x$ctss)) %>% 
-    map2_dfr(., names(.), ~tibble(visit = .y, p_value = .x$p.value))
+    map(~shapiro_test(.x$ctss)) %>% 
+    map2_dfr(., names(.), ~mutate(.x, visit = .y))
   
   eda$ctss_normality$time_sev_plot_ident <- radio$long %>% 
     dlply(.(visit, severity)) %>% 
@@ -143,8 +143,8 @@
   
   eda$ctss_normality$time_test_log <- radio$long %>% 
     dlply(.(visit)) %>% 
-    map(~shapiro.test(log(.x$ctss + 1))) %>% 
-    map2_dfr(., names(.), ~tibble(visit = .y, p_value = .x$p.value))
+    map(~shapiro_test(log(.x$ctss + 1))) %>% 
+    map2_dfr(., names(.), ~mutate(.x, visit = .y))
   
   eda$ctss_normality$time_plot_log <- radio$long %>% 
     dlply(.(visit)) %>% 
@@ -156,14 +156,27 @@
   
   eda$ctss_normality$time_sev_test_log <- radio$long %>% 
     dlply(.(visit, severity)) %>% 
-    map(~shapiro.test(log(.x$ctss + 1))) %>% 
-    map2_dfr(., names(.), ~tibble(visit = .y, p_value = .x$p.value))
+    map(~shapiro_test(log(.x$ctss + 1))) %>% 
+    map2_dfr(., names(.), ~mutate(.x, visit = .y))
   
   eda$ctss_normality$time_sev_plot_log <- radio$long %>% 
     dlply(.(visit, severity)) %>% 
     map(~ggplot(.x, aes(sample = log(ctss + 1))) + 
           geom_qq() + 
           geom_qq_line())
+
[…]

### 2021-12-27 analysis scripts/risk_score.R

@@ -133,6 +133,23 @@
   
   plan('sequential')
   
+  ## model diagnostic tables, normality and variance homogeneity testing
+  
+  risk$multivariate_resids <- risk$multivariate_models %>% 
+    map(~.x$finalModel) %>% 
+    map(get_qc_tbl)
+
+  risk$multivariate_normality <- risk$multivariate_resids %>% 
+    map(~.x$.resid) %>% 
+    map(shapiro_test) %>% 
+    map2_dfr(., names(.), ~mutate(.x, variable = .y))
+  
+  ## diagnostic plots
+  
+  risk$multivariate_diagnostic <- risk$multivariate_models %>% 
+    map(~.x$finalModel) %>% 
+    map(get_qc_plots)
+
   ## final models, inference
   
   risk$multivariate_summary <- risk$multivariate_models %>% 

### 2021-12-27 analysis scripts/eda.R
Add files via upload

@@ -14,12 +14,12 @@
   
   ## analyses
 
-  eda$baseline$analyses_cohort <- radio$long_expl_var %>% 
+  eda$baseline$analyses_cohort <- c('age', radio$long_expl_var) %>% 
     map(analyze_feature, 
         inp_tbl = rbind(mutate(radio$clear, dummy_var = factor('dummy1')), 
                         mutate(radio$clear, dummy_var = factor('dummy2'))), 
         split_var = 'dummy_var') %>% 
-    set_names(radio$long_expl_var)
+    set_names(c('age', radio$long_expl_var))
 
   ## summary table
   
@@ -35,18 +35,18 @@
   
   ## analyses
   
-  eda$baseline$analyses_severity <- radio$long_expl_var %>% 
+  eda$baseline$analyses_severity <- c('age', radio$long_expl_var) %>% 
     map(analyze_feature, 
         inp_tbl = radio$clear, 
         split_var = 'severity') %>% 
-    set_names(radio$long_expl_var)
+    set_names(c('age', radio$long_expl_var))
   
   ## summary tables, for age: ANOVA (homogeneous variances and normality, see the analysis object)
   
   eda$baseline$summary_severity <- eda$baseline$analyses_severity %>% 
     map_dfr(get_feature_summary) %>% 
     filter(variable != 'severity') %>% 
-    mutate(p_value = ifelse(is.na(p_chi), p_param, p_chi)) %>% 
+    mutate(p_value = ifelse(is.na(p_chi), p_non_param, p_chi)) %>% 
     add_stat_info %>% 
     select(variable, label, mild, moderate, severe, critical, test, p_value, significance, p_adj)
   

### 2021-12-29 paper/markdown/figures.Rmd
Add files via upload

@@
 ```{r tab-baseline-severity, tab.cap = 'Baseline characteristic of the acute COVID-19 severity subsets.
-Statistical significance of the differences between the severity subsets was assessed with $\\chi^2$ test.'}
+Statistical significance of the differences between the severity subsets was assessed with $\\chi^2$ (categorical variables) or Mann-Whitney U test (age).'}
 
@@
   width(2:5, width = 3.9, unit = 'cm') %>% 
-  width(6, width = 2.5, unit = 'cm')
+  width(6, width = 2.5, unit = 'cm') %>% 
+  merge_v(1)
 
@@
   width(3:6, width = 3.2, unit = 'cm') %>% 
-  width(7, width = 2.5, unit = 'cm')
+  width(7, width = 2.5, unit = 'cm') %>% 
+  merge_v(1)
 
@@
   width(6, width = 2.5, unit = 'cm')
+
+```
+
+\newpage
+
+```{r tab-opacity, tab.cap = 'Percentages of opacity and high opacity at the consecutive visits in the entire cohort and the acute COVID-19 severity subsets.
+Statistical significance of the differences between the visits was assessed with Kruskal-Wallis test.'}
+
+paper_tables$opacity %>% 
+  flextable() %>%
+  width(1, width = 1.8, unit = 'cm') %>% 
+  width(2, width = 1.8, unit = 'cm') %>% 
+  width(3:6, width = 3.5, unit = 'cm') %>% 
+  width(7, width = 2.5, unit = 'cm') %>% 
+  merge_v(1)
 
@@
 
-__(A)__ Risk factors of developing PILI at the one-year follow-up visit were identified by a series of logistic models.
+__(A)__ Risk factors of developing any lung CT abnormality at the one-year follow-up visit were identified by a series of logistic models.
 Odds ratio (OR) significance was determined by Wald Z test.
@@
 
-__(B)__ Independent risk factors of PILI were identified by multi-parameter logistic modeling with backward elimination.
+__(B)__ Independent risk factors of lung CT abnormality were identified by multi-parameter logistic modeling with backward elimination.
 OR significance was determined by Wald Z test.
@@
 OR with 95$%$ confidence intervals are presented in a Forest plot.
-Area under the curve (AUC), sensitivity (Se) and specificity (Sp) for detection of PILI in CV are indicated in the plot caption.
+Area under the curve (AUC), sensitivity (Se) and specificity (Sp) for detection of lung CT abnormality in CV are indicated in the plot caption.
 Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.
 
-BMI: body mass index, PKY: packages per year.
+Orange: positive correlation, blue: negative correlation, gray: not significant, BMI: body mass index, PKY: packages per year, ref.: reference.
 
@@
 
-BMI: body mass index, PKY: packages per year.
+Orange: positive correlation, blue: negative correlation, gray: not significant, BMI: body mass index, PKY: packages per year, ref.: reference.
+
+\newpage
+
+# Supplementary Figures
+
+\beginsupplement
+
+```{r fig-ct-opacity, fig.width = mm_inch(suppl_figures$ct_opacity$w), fig.height = mm_inch(suppl_figures$ct_opacity$h), dpi = 600, fig.cap = 'Resolution kinetic of lung CT opacity.'}
+
+suppl_figures$ct_opacity$plot
+
+```
+
+__Figure \@ref(fig:fig-ct-opacity). Resolution kinetic of lung CT opacity.__
+
+Risk of any lung CT opacity at the consecutive time points was modeled by mixed-effect logistic regression (random effect: individual, fixed effect: time post COVID-19) in the entire cohort and the acute COVID-19 severity subsets.
+Significance of the fixed time term was determined by likelihood ratio test (LRT) versus the respective random effect-only null model.
+Plots display percents of individuals with CT opacity in the analyzed collectives.
+$\chi$ statistic, degrees of freedom and p values of the LRT test are presented in the plot captions.
+Numbers of individuals with the complete set of consecutive CT scans are shown under the plots.
+
+\newpage
+
+```{r fig-opacity-kinetic, fig.width = mm_inch(suppl_figures$opacity_kinetic$w), fig.height = mm_inch(suppl_figures$opacity_kinetic$h), dpi = 600, fig.cap = 'Kinetic of the lung opacity.'}
+
+suppl_figures$opacity_kinetic$plot
+
+```
+
+__Figure \@ref(fig:fig-opacity-kinetic). Kinetic of the lung opacity.__
+
+Kinetic of the lung opacity at the consecutive time points was investigated by Friedman test (grouping by the individual) in the entire cohort and the acute COVID-19 severity subsets.
+The effect size was determined by Kendall W test, differences between particular time points were compared by Wilcoxon test.
+Plots display individual opacity trajectories as thin gray lines, thick colored lines represent medians, interquartile ranges are presented as colored regions.
+$\chi$ statistic, degrees of freedom and p values of the Friedman test and the Kendall W statistic with 95$%$ confidence intervals are presented in the plot captions.
+Numbers of individuals with the complete set of consecutive CT scans are shown under the plots.
+
+\newpage
+
+```{r fig-opacity-risk, fig.width = mm_inch(suppl_figures$opacity_risk$w), fig.height = mm_inch(suppl_figures$opacity_risk$h), dpi = 600, fig.cap = 'Risk of developing lung opacity at the one-year post-COVID-19 follow-up.'}
+
+suppl_figures$opacity_risk$plot
+
+```
+
+__Figure \@ref(fig:fig-opacity-risk). Risk of developing lung opacity at the one-year post-COVID-19 follow-up.__
+
+__(A)__ Risk factors of developing opacity at the one-year follow-up visit were identified by a series of logistic models.
+Odds ratio (OR) significance was determined by Wald Z test.
+OR with 95$%$ confidence intervals are presented in a Forest plot.
+Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.
+
+__(B)__ Independent risk factors of opacity were identified by multi-parameter logistic modeling with backward elimination.
+OR significance was determined by Wald Z test.
+Model prediction was verified by 20-fold cross-validation (CV) and receiver-operating characteristic (ROC).
+OR with 95$%$ confidence intervals are presented in a Forest plot.
+Area under the curve (AUC), sensitivity (Se) and specificity (Sp) for detection of opacity in CV are indicated in the plot caption.
+Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.
+
+Orange: positive correlation, blue: negative correlation, gray: not significant, BMI: body mass index, PKY: packages per year, ref.: reference.
+
+\newpage
+
+```{r fig-opacity-rater, fig.width = mm_inch(suppl_figures$any_abnormality_rater$w), fig.height = mm_inch(suppl_figures$any_abnormality_rater$h), dpi = 600, fig.cap = 'Performance of detection of lung CT abnormality by automated opacity assessment.'}
+
+suppl_figures$any_abnormality_rater$plot
+
+```
+
+__Figure \@ref(fig:fig-opacity-rater). Performance of detection of lung CT abnormality by automated opacity assessment.__
+
+Lung CT abnormality were defined either as CTSS $\geq$ 1 0 or as opacity $\geq$ 1$%$ lung volume.
+Performance of the opacity at detecting CT abnormalities defined by CTSS was investigated at the consecutive follow-ups by receiver-operating characteristic (ROC) and Cohen's $\kappa$ statistic.
+Numbers of complete observations and CT abnormalities (events) are indicated under the plots.
+
+__(A)__ ROC curves annotated with sensitivity (Se), specificity (Sp) and area under the curve (AUC) values with 95$%$ confidence intervals.
+
+__(B)__ $\kappa$ with 95$%$ confidence intervals presented in a forest plot.
+
+\newpage
+
+```{r fig-ctss-opacity, fig.width = mm_inch(suppl_figures$ctss_opacity$w), fig.height = mm_inch(suppl_figures$ctss_opacity$h), dpi = 600, fig.cap = 'Correlation of CTSS and lung opacity.'}
+
+suppl_figures$ctss_opacity$plot
+
+```
+
+__Figure \@ref(fig:fig-ctss-opacity). Correlation of CTSS and lung opacity.__
+
+Correlation of CTSS and lung opacity at the consecutive follow-ups was investigated with Spearman test.
+Points represent single observations, blue lines with gray regions represent fitted linear trend with 95$%$ confidence interval.
+Correlation coefficients (r), p values and numbers of complete observations are indicated in the plot captions.

### 2021-12-29 analysis scripts/distribution.R

@@ -1,0 +1,140 @@
+# This script performs distribution checks of the numeric outcome variables
+
+  insert_head()
+  
+# container list -----
+  
+  distr <- list()
+  
+# globals -----
+  
+  insert_msg('Globals setup')
+  
+  distr$responses <- c('ctss', 'perc_opac')
+  
+# Normality of the CT severity score, percent opacity and percent high opacity at the time points -----
+  
+  insert_msg('CT severity score normality check')
+  
+  ## normality in the visit groups
+  
+  distr$ctss_normality$time_test_ident <- radio$long %>% 
+    dlply(.(visit)) %>% 
+    map(~shapiro_test(.x, ctss, perc_opac)) %>% 
+    map2_dfr(., names(.), ~mutate(.x, visit = .y))
+
+  distr$ctss_normality$time_plot_ident$ctss <- radio$long %>% 
+    dlply(.(visit)) %>% 
+    map(~ggplot(.x, aes(sample = ctss)) + 
+          geom_qq() + 
+          geom_qq_line())
+  
+  distr$ctss_normality$time_plot_ident$perc_opac <- radio$long %>% 
+    dlply(.(visit)) %>% 
+    map(~ggplot(.x, aes(sample = perc_opac)) + 
+          geom_qq() + 
+          geom_qq_line())
+  
+  ## normality in the time and time:severity groups
+  
+  distr$ctss_normality$time_sev_test_ident <- radio$long %>% 
+    dlply(.(visit, severity)) %>% 
+    map(~shapiro_test(.x, ctss, perc_opac)) %>% 
+    map2_dfr(., names(.), ~mutate(.x, visit = .y))
+  
+  distr$ctss_normality$time_sev_plot_ident$ctss <- radio$long %>% 
+    dlply(.(visit, severity)) %>% 
+    map(~ggplot(.x, aes(sample = ctss)) + 
+          geom_qq() + 
+          geom_qq_line())
+  
+  distr$ctss_normality$time_sev_plot_ident$perc_opac <- radio$long %>% 
+    dlply(.(visit, severity)) %>% 
+    map(~ggplot(.x, aes(sample = perc_opac)) + 
+          geom_qq() + 
+          geom_qq_line())
+  
+  ## normality in the visit groups
+  
+  distr$ctss_normality$time_test_log <- radio$long %>% 
+    dlply(.(visit)) %>% 
+    map(mutate, 
+        log_ctss = log(ctss + 1), 
+        log_perc_opac = log(perc_opac + 1)) %>% 
[…]

### 2021-12-29 analysis scripts/eda.R

@@ -109,97 +109,31 @@
     add_stat_info %>%
     select(visit, variable, label, mild, moderate, severe, critical, test, p_value, significance, p_adj)
   
-# Normality of the CT severity score at the time points -----
-  
-  insert_msg('CT severity score normality check')
-  
-  ## normality in the visit groups
-  
-  eda$ctss_normality$time_test_ident <- radio$long %>% 
-    dlply(.(visit)) %>% 
-    map(~shapiro_test(.x$ctss)) %>% 
-    map2_dfr(., names(.), ~mutate(.x, visit = .y))
-  
-  eda$ctss_normality$time_plot_ident <- radio$long %>% 
-    dlply(.(visit)) %>% 
-    map(~ggplot(.x, aes(sample = ctss)) + 
-          geom_qq() + 
-          geom_qq_line())
-  
-  ## normality in the time and time:severity groups
-  
-  eda$ctss_normality$time_sev_test_ident <- radio$long %>% 
-    dlply(.(visit, severity)) %>% 
-    map(~shapiro_test(.x$ctss)) %>% 
-    map2_dfr(., names(.), ~mutate(.x, visit = .y))
-  
-  eda$ctss_normality$time_sev_plot_ident <- radio$long %>% 
-    dlply(.(visit, severity)) %>% 
-    map(~ggplot(.x, aes(sample = ctss)) + 
-          geom_qq() + 
-          geom_qq_line())
-  
-  ## normality in the visit groups
-  
-  eda$ctss_normality$time_test_log <- radio$long %>% 
-    dlply(.(visit)) %>% 
-    map(~shapiro_test(log(.x$ctss + 1))) %>% 
-    map2_dfr(., names(.), ~mutate(.x, visit = .y))
-  
-  eda$ctss_normality$time_plot_log <- radio$long %>% 
-    dlply(.(visit)) %>% 
-    map(~ggplot(.x, aes(sample = log(ctss + 1))) + 
-          geom_qq() + 
-          geom_qq_line())
-  
-  ## normality of the log CTSS in the time and time:severity groups
-  
-  eda$ctss_normality$time_sev_test_log <- radio$long %>% 
-    dlply(.(visit, severity)) %>% 
-    map(~shapiro_test(log(.x$ctss + 1))) %>% 
-    map2_dfr(., names(.), ~mutate(.x, visit = .y))
-  
[…]

### 2021-12-29 analysis scripts/interrater.R

@@ -1,0 +1,191 @@
+# This script compares the performance of the manual CTSS and automated opacity detection at the consecutive time points
+
+  insert_head()
+  
+# container list ----
+  
+  rater <- list()
+
+# inter-rater comparison, any abnormalities by CTSS and opacity ----
+  
+  insert_msg('Any abnormalities')
+
+  ## table objects
+    
+  rater$any_ct$tables <- radio$long %>% 
+    rater_tbl(ctss_any, opacity) %>% 
+    map(~table(.x[c('ctss_any', 'opacity')]))
+  
+  ## n numbers
+  
+  rater$any_ct$n_numbers <- map2(map(rater$any_ct$tables, sum), 
+                                 map(rater$any_ct$tables, ~sum(.x[2, ])), 
+                                 ~paste0('total: n = ', .x, ', events: n = ', .y))
+  
+  rater$any_ct$n_tag <- map2_chr(globals$visit_labels, 
+                                 rater$any_ct$n_numbers, 
+                                 paste, sep = '\n')
+  
+  ## Kappa calculation
+  
+  rater$any_ct$kappa <- rater$any_ct$tables %>% 
+    map(Kappa_pipe) %>% 
+    map2_dfr(., names(.), ~mutate(.x, visit = .y)) %>% 
+    mutate(visit = factor(visit, levels = rev(visit)))
+  
+  ## ROC
+  
+  rater$any_ct$cutpoint <- radio$long %>% 
+    rater_tbl(ctss_any, opacity) %>% 
+    map(~map_dfc(.x, ~as.numeric(.x) - 1)) %>% 
+    map(find_optimal_cutoff, 
+        status_variable = 'ctss_any', 
+        marker = 'opacity')
+  
+  rater$any_ct$roc_stats <- rater$any_ct$cutpoint %>% 
+    map(get_optimal_cutpoint_coefs) %>% 
+    map(unlist, recursive = FALSE) %>% 
+    map(~.x[2:3]) %>%
+    as_tibble %>% 
+    t %>% 
+    as.data.frame %>% 
+    set_names(c('Se', 'Sp')) %>% 
+    rownames_to_column('visit')
+  
+  rater$any_ct$roc_stats <- rater$any_ct$cutpoint %>% 
+    map_dfr(get_auc) %>% 
+    cbind(rater$any_ct$roc_stats, .) %>% 
+    as_tibble %>% 
+    mutate(plot_annotation = paste0('Se = ', signif(Se, 2), 
[…]

### 2021-12-29 analysis scripts/kinetic.R

@@ -18,60 +18,101 @@
     dlply(.(severity)) %>% 
     c(list(cohort = kinetic$analysis_tbl), .)
   
-# CTSS kinetics -----
+# CTSS, opacity and hig opacity percentage kinetics -----
   
   insert_msg('CTSS kinetic')
   
   ## test
   
-  kinetic$ctss$analysis <- kinetic$analysis_tbl %>% 
-    map(~friedman_test(ctss ~ visit|ID, data = .x))
-  
+  kinetic$numeric$analysis <- list(ctss = ctss ~ visit|ID, 
+                                   perc_opac = perc_opac ~ visit|ID, 
+                                   perc_hiopac = perc_hiopac ~ visit|ID) %>% 
+    map(function(outcome) kinetic$analysis_tbl %>% 
+          map(friedman_test, 
+                  formula = outcome) %>% 
+          map2_dfr(., names(.), ~mutate(.x, subset = .y)))
+
   ## summary
   
-  kinetic$ctss$summary <- kinetic$ctss$analysis %>% 
-    map2_dfr(., names(.), ~mutate(.x, subset = .y, p_value = p)) %>% 
-    mutate(p_adj = p.adjust(p_value, 'BH'), 
-           plot_caption = paste0('\u03C7 = ', signif(statistic, 2), 
-                                 ', df = 3, p = ', signif(p_value, 2)))
+  kinetic$numeric$summary <- kinetic$numeric$analysis %>% 
+    map(mutate, 
+        p_value = p, 
+        p_adj = p.adjust(p_value, 'BH'), 
+        plot_caption = paste0('\u03C7 = ', signif(statistic, 2), 
+                              ', df = 3, ', format_p(p_value, 2)))
   
   ## effect size determined by Kendall W, updating the plot_caption
   
   plan('multisession')
   
-  kinetic$ctss$effsize <- kinetic$analysis_tbl %>% 
-    future_map(~friedman_effsize(ctss ~ visit|ID, 
-                                 data = .x, 
-                                 ci = TRUE, 
-                                 ci.type = 'bca'), 
-               .options = furrr_options(seed = TRUE)) %>% 
-    map2_dfr(., names(.), ~mutate(.x, subset = .y))
-  
-  kinetic$ctss$summary <- kinetic$ctss$summary %>% 
-    mutate(plot_caption = list(x = plot_caption, 
-                               y = kinetic$ctss$effsize$effsize, 
[…]

### 2021-12-29 analysis scripts/risk_score.R

@@ -1,7 +1,7 @@
 # This script models risk of any, moderate-severe, severe, ggo, reticulation, consolidations and bronchial dilatation 
 # (logistic regression) as well as of the abnormality grading at the 1-year follow-up. Explanatory variables
 # are> sex, age, severity, bin, smoking parameters as well as initial CT abnormality severity (CTSS with the cutoffs 5 and 10)
-# Multi-paramater models are constructed by backward elimination and validated by 20-fold CV
+# Multi-parameter models are constructed by backward elimination and validated by 20-fold CV
 
   insert_head()
   
@@ -18,17 +18,21 @@
   
   risk$outcomes <-  c('ggo_fup4', 
                       'retic_fup4', 
-                      'pili_fup4', 
                       'ctss_any_fup4', 
-                      'ctss_fup4')
+                      'opacity_fup4', 
+                      'ctss_fup4', 
+                      'log_perc_opac_fup4')
   
   risk$outcome_forms <- risk$outcomes %>% 
     map(~paste(.x, paste(risk$explanatory_vars, collapse = '+'), sep = '~')) %>% 
     map(as.formula)
   
-  risk$families <- c(rep('binomial', 4), 'poisson')
+  risk$families <- c(rep('binomial', 4), 
+                     'poisson', 
+                     'gaussian')
   
-  risk$analysis_tbl <- radio$clear
+  risk$analysis_tbl <- radio$clear %>% 
+    mutate(log_perc_opac_fup4 = log(perc_opac_fup4 + 1))
   
   for(i in risk$outcomes) {
     
@@ -97,13 +101,14 @@
   risk$univariate_plots <- list(summary_tbl = risk$univariate_summary, 
                                 plot_title = c('GGO, 1-year FUP', 
                                                'Reticulations, 1-year FUP', 
-                                               'PILI, 1-year FUP', 
-                                               'Any CT abnormality, 1-year FUP', 
-                                               'CTSS, 1-year FUP'), 
+                                               'CT abnormality, 1-year FUP', 
+                                               'Opacity, 1-year FUP', 
+                                               'CTSS, 1-year FUP', 
+                                               '% opacity, 1-year FUP'), 
                                 plot_subtitle = c(rep('Logistic regression', 4),  
-                                                      'Poisson regression'), 
-                                hide_baseline_est = c(rep(TRUE, 4), FALSE), 
[…]

### 2021-12-29 data_analysis.R

@@ -5,11 +5,12 @@
 # tools -----
   
   c('./tools/sys_tools.R', 
-    './tools/project_tools.R', 
     './tools/project_globals.R', 
     './tools/counting_tools.R', 
     './tools/modeling_tools.R', 
-    './tools/lm_qc_tools.R') %>% 
+    './tools/lm_qc_tools.R', 
+    './tools/roc_toolbox.R', 
+    './tools/project_tools.R') %>% 
     walk(source)
   
   library(DescTools)
@@ -17,12 +18,15 @@
   library(ggpubr)
   library(furrr)
   library(caret)
+  library(vcd)
   
 # analysis scripts -----
   
-  c('./analysis scripts/eda.R', 
-    './analysis scripts/kinetic.R', 
-    './analysis scripts/risk_score.R') %>% 
+  c('./analysis scripts/distribution.R', ## distribution checks
+    './analysis scripts/eda.R', ## explorative data analysis
+    './analysis scripts/kinetic.R', ## kinetic analysis
+    './analysis scripts/risk_score.R', ## risk and score modeling
+    './analysis scripts/interrater.R') %>% ## inter-rater comparison of the CTSS and opacity
     walk(source)
   
 # END -----

### 2021-12-29 data_import.R

@@ -23,7 +23,7 @@
   
   insert_msg('Reading the SPSS data file')
   
-  radio$raw <- read.spss(file = './data/20211216PILI.sav', 
+  radio$raw <- read.spss(file = './data/20211228PILI.sav', 
                           to.data.frame = TRUE) %>% 
     as_tibble %>% 
     mutate(ID = rm_spaces(RAD_ID, 'character'))
@@ -78,8 +78,8 @@
            ctss_any_fup4 = factor(ifelse(ctss_fup4 > 0, 'yes', 'no')), 
            ctss_mod_severe_fup4 = factor(ifelse(ctss_fup4 > 5, 'yes', 'no')), 
            ctss_severe_fup4 = factor(ifelse(ctss_fup4 > 10, 'yes', 'no')), 
-           pili_fup4 = ifelse(pili_fup4 == 'no abnormalities', 'no', 'yes'), 
-           pili_fup4 = factor(pili_fup4)) %>% 
+           opacity_fup4 = factor(ifelse(perc_opac_fup4 > 0, 'yes', 'no')), 
+           hiopacity_fup4 = factor(ifelse(perc_hiopac_fup4 > 0, 'yes', 'no'))) %>% 
     filter(!is.na(date_fup4)) %>% 
     map_dfr(function(x) if(is.factor(x)) droplevels(x) else x)
   
@@ -95,10 +95,12 @@
   
   radio$long_outcome_var <- radio$vars %>% 
     filter(modeling == 'long_outcome') %>% 
+    arrange(order) %>% 
     .$new_var
   
   radio$long_expl_var <- radio$vars %>% 
     filter(modeling == 'explanatory') %>% 
+    arrange(order) %>% 
     .$new_var
   
   radio$long <- radio$long_outcome_var %>% 
@@ -138,16 +140,20 @@
            ctss_class = factor(ctss_class, levels = levels(radio$clear$ctss_class_fup1)), 
            ctss_any = factor(ifelse(ctss > 0, 'yes', 'no')), 
            ctss_mod_severe = factor(ifelse(ctss > 5, 'yes', 'no')), 
-           ctss_severe = factor(ifelse(ctss > 10, 'yes', 'no')))
+           ctss_severe = factor(ifelse(ctss > 10, 'yes', 'no')), 
+           opacity = factor(ifelse(perc_opac > 0, 'yes', 'no')), 
+           hiopacity = factor(ifelse(perc_hiopac > 0, 'yes', 'no')))
 
[…]

### 2021-12-29 paper scripts/paper_figures.R

@@ -2,9 +2,10 @@
 
   insert_head()
 
-# container list -----
+# container lists -----
 
   paper_figures <- list()
+  suppl_figures <- list()
   
 # Figure 1: Consort plot - placeholder, provided by the Author team ----
   
@@ -59,11 +60,11 @@
   
   insert_msg('Figure 3: kinetic of the CTSS')
   
-  paper_figures$ctss_kinetic <- kinetic$ctss$plots[c('cohort',
-                                                     'mild', 
-                                                     'moderate', 
-                                                     'severe', 
-                                                     'critical')] %>% 
+  paper_figures$ctss_kinetic <- kinetic$numeric$plots_ctss[c('cohort',
+                                                             'mild', 
+                                                             'moderate', 
+                                                             'severe', 
+                                                             'critical')] %>% 
     plot_grid(plotlist = ., 
               ncol = 2, 
               align = 'hv') %>% 
@@ -71,22 +72,22 @@
                      w = 180, 
                      h = 210)
   
-# Figure 5: PILI modeling -----
+# Figure 4: Any CT abnormality modeling -----
   
-  insert_msg('Figure 4: PILI modeling')
+  insert_msg('Figure 4: Any CT abnormality modeling')
   
-  paper_figures$pili_modeling <- plot_grid(risk$univariate_plots$pili_fup4, 
-                                           risk$multivariate_plots$pili_fup4, 
+  paper_figures$pili_modeling <- plot_grid(risk$univariate_plots$ctss_any_fup4, 
+                                           risk$multivariate_plots$ctss_any_fup4, 
                                            nrow = 2, 
                                            align = 'hv', 
-                                           rel_heights = c(0.6, 0.4), 
+                                           rel_heights = c(0.63, 0.37), 
                                            labels = LETTERS, 
                                            label_size = 10) %>% 
     as_figure_object(figure_label = 'figure_4_pili_modeling', 
                      w = 180, 
                      h = 210)
   
-# Figure 6: CTSS modeling ------
+# Figure 5: CTSS modeling ------
   
   insert_msg('Figure 5: CTSS modeling')
   
@@ -94,13 +95,106 @@
                                            risk$multivariate_plots$ctss_fup4, 
                                            nrow = 2, 
                                            align = 'hv', 
-                                           rel_heights = c(0.6, 0.4), 
+                                           rel_heights = c(0.64, 0.36), 
[…]

### 2021-12-29 paper scripts/paper_tables.R

@@ -58,10 +58,10 @@
           starts_with('fup'), 
           significance) %>% 
    set_names(c('Variable', 
-               '60-day FUP', 
-               '100-day FUP', 
-               '180-day FUP', 
-               '360-day FUP', 
+               '2-month FUP', 
+               '3-month FUP', 
+               '6-month FUP', 
+               '1-year FUP', 
                'Significance'))
  
 # Table 6: CT abnormalities at the visits, severity ----
@@ -75,10 +75,10 @@
           significance) %>% 
    set_names(c('Severity', 
                'Variable', 
-               '60-day FUP', 
-               '100-day FUP', 
-               '180-day FUP', 
-               '360-day FUP', 
+               '2-month FUP', 
+               '3-month FUP', 
+               '6-month FUP', 
+               '1-year FUP', 
                'Significance'))
  
 # Table 6: CT abnormalities between the severity groups -----
@@ -106,16 +106,36 @@
  
  insert_msg('Table 7: CTSS at the consecutive visits')
  
- paper_tables$ctss <- eda$ctss$summary %>% 
+ paper_tables$ctss <- eda$numeric$summary$ctss %>% 
    select(subset, 
           starts_with('fup'), 
           significance) %>% 
    set_names(c('Severity',  
-               '60-day FUP', 
-               '100-day FUP', 
-               '180-day FUP', 
-               '360-day FUP', 
+               '2-month FUP', 
+               '3-month FUP', 
+               '6-month FUP', 
+               '1-year FUP', 
                'Significance'))
+ 
+# Table 8: percent opacity and high opacity -----
+ 
+ insert_msg('Table 8: Percent opacity and high opacity')
+ 
+ paper_tables$opacity <- eda$numeric$summary[c('perc_opac', 'perc_hiopac')] %>% 
+    map2_dfr(., c('Opacity, % lung volume', 
+                  'High opacity, % lung volume'),  
+             ~mutate(.x, variable = .y)) %>% 
+    select(variable, 
+           subset, 
+           starts_with('fup'), 
+           significance) %>% 
+    set_names(c('Variable', 
+                'Severity',  
+                '2-month FUP', 
+                '3-month FUP', 
+                '6-month FUP', 
+                '1-year FUP', 
+                'Significance'))
  
 # Saving the tables as an excel file ----
  

### 2021-12-29 render_paper.R

@@ -16,7 +16,8 @@
   insert_msg('Sourcing the paper scripts')
   
   c('./paper scripts/paper_tables.R', 
-    './paper scripts/paper_figures.R') %>% 
+    './paper scripts/paper_figures.R', 
+    './paper scripts/deploy_paper.R') %>% 
     walk(source)
   
 # END -----

### 2021-12-29 tools/counting_tools.R

@@ -838,9 +838,11 @@
   
 # making a summary table -----
   
-  get_feature_summary <- function(analysis_object, label = NA, signif_digits = 3) {
+  get_feature_summary <- function(analysis_object, label = NA, signif_digits = 2, short_numerics = FALSE) {
     
     ## makes a table with cohort characteristics which may be used in your paper
+    
+    short_fun <- if(short_numerics) short_number else signif
     
     if(class(analysis_object) == 'list') {
       
@@ -858,21 +860,21 @@
       desc_stats <- analysis_object$stat_tables %>% 
         map(mutate, 
             mean_cell = paste('mean(SD) = ', 
-                              signif(mean, signif_digits), 
+                              short_fun(mean, signif_digits), 
                               ' (', 
-                              signif(sd, signif_digits), 
+                              short_fun(sd, signif_digits), 
                               ')', sep = ''), 
             median_cell = paste('median(IQR) = ', 
-                                signif(median, signif_digits), 
+                                short_fun(median, signif_digits), 
                                 ' (', 
-                                signif(perc25, signif_digits), 
+                                short_fun(perc25, signif_digits), 
                                 ' - ', 
-                                signif(perc75, signif_digits), 
+                                short_fun(perc75, signif_digits), 
                                 ')', sep = ''), 
             min_max_cell = paste('range = ', 
-                                 signif(min, signif_digits), 
+                                 short_fun(min, signif_digits), 
                                  ' - ', 
-                                 signif(max, signif_digits), 
+                                 short_fun(max, signif_digits), 
                                  sep = ''), 
             tbl_cell = paste(mean_cell, 
                              median_cell, 
@@ -1036,6 +1038,16 @@
     
     return(paste('p =', 
                  signif(p_value, signif_digits)))
+    
+  }
+  
+  short_number <- function(vec, signif_digits = 2) {
+    
+    stopifnot(is.numeric(vec))
+    
+    signif(vec, signif_digits) %>% 
+      as.character %>% 
+      stri_replace(regex = '^0.', replacement = '.')
     
   }
   

### 2021-12-29 tools/project_globals.R

@@ -52,5 +52,15 @@
                              severe = 'Severe COVID-19', 
                              critical = 'Critical COVID-19')
   
+  globals$visit_labels <- c('fup1' = '2-month', 
+                            'fup2' = '3-month', 
+                            'fup3' = '6-month', 
+                            'fup4' = '1-year')
+  
+  globals$visit_colors <- c('fup1' = 'firebrick3', 
+                            'fup2' = 'cornsilk4', 
+                            'fup3' = 'darkolivegreen4', 
+                            'fup4' = 'steelblue')
+  
 # END -----
   

### 2021-12-29 tools/project_tools.R

@@ -5,7 +5,7 @@
   require(rlang)
   require(stringi)
   require(foreign)
-
+  require(vcd)
   
 # data import and transformation -----
   
@@ -247,9 +247,7 @@
     
     data_frame %>% 
       mutate(p_adj = p.adjust(p_value, 'BH'), 
-             significance = ifelse(.data[[signif_var]] < 0.05, 
-                                   paste('p =', signif(p_value, 2)), 
-                                   paste0('ns (p = ', signif(p_value, 2), ')')), 
+             significance = format_p(.data[[signif_var]]), 
              label = translate_var(variable), 
              test = ifelse(class == 'factor', 'Chi squared', 'one-way ANOVA')) %>% 
       map_dfc(function(x) if(is.character(x)) stri_replace(x, regex = '^.*\\nyes:\\s{1}', replacement = '') else x)
@@ -259,6 +257,8 @@
 # kinetic plotting -----
   
   plot_ctss <- function(data, 
+                        plot_var = 'ctss', 
+                        y_lab = 'CTSS', 
                         plot_title = NULL, 
                         plot_subtitle = NULL, 
                         fill_color = 'firebrick') {
@@ -271,14 +271,14 @@
     med_tbl <- data %>% 
       dlply(.(visit)) %>% 
       map_dfr(analyze_feature, 
-              variable = 'ctss') %>% 
+              variable = plot_var) %>% 
       mutate(time_months = c(2, 3, 6, 12))
     
     ## plot
     
     data %>% 
       ggplot(aes(x = time_months, 
-                 y = ctss)) + 
+                 y = .data[[plot_var]])) + 
       geom_ribbon(data = med_tbl, 
                   aes(y = median, 
                       ymin = perc25, 
@@ -297,7 +297,7 @@
            subtitle = plot_subtitle, 
            tag = paste('\nn =', med_tbl$n_complete[1]), 
            x = 'Months post COVID-19', 
-           y = 'CTSS')
+           y = y_lab)
     
   }
   
@@ -351,13 +351,16 @@
   add_ctss_post_p <- function(plot, 
                               post_hoc_results, 
                               p_var = 'p', 
-                              hoffset = 0.15) {
+                              hoffset = 0.15, 
+                              line_y = 21, 
+                              voffset = 2, 
+                              text_offset = 0.35) {
     
     ## adds the results of sequential post-hoc testing to the ctss plot
     
[…]

### 2021-12-29 tools/roc_toolbox.R

@@ -1,0 +1,111 @@
+# A set of useful functions for ROC plotting and analysis
+
+# libraries -----
+
+  library(OptimalCutpoints)
+  library(plotROC)
+
+# globals ----
+
+  ## a common theme for ROC
+  
+    roc_theme <-  theme(legend.text = element_text(size = 8, face = 'plain', color = 'black'), 
+                        legend.title = element_text(size = 8, face = 'plain', color = 'black'), 
+                        axis.text = element_text(size = 8, face = 'plain', color = 'black'), 
+                        axis.title = element_text(size = 8, face = 'plain', color = 'black'))
+    
+    
+# ROC analysis and plotting toolbox ----
+    
+    find_optimal_cutoff <- function(inp_table, status_variable, marker, tag.healthy = 0, method = 'Youden', ...) {
+      
+      ## returns an output of the optimal.cutpoints from the package OptimalCutpoints
+      
+      analysis_table <- inp_table %>% 
+        data.frame
+      
+      output <- optimal.cutpoints(status = status_variable, X = marker, data = analysis_table, 
+                                  tag.healthy = tag.healthy, method = 'Youden', ...)
+      
+      return(output)
+      
+    }
+    
+    get_optimal_cutpoint_coefs <- function(optimal_cutpoint_object) {
+      
+      ## extracts coefficient values from the optimal_cutpoint_object generated e.g. by 
+      ## the find_optimal_cutoff function
+      
+      output <- optimal_cutpoint_object %>% 
+        summary
+      
+      output <- output$p.table$Global$Youden
+      
+      return(output)
+      
+    }
+    
+    get_auc <- function(optimal_cutpoint_object) {
+      
+      require(stringi)
+      
+      ## extracts AUC with its confidence intetervals from the optimal_cutpoint_object generated e.g. by 
+      ## the find_optimal_cutoff function
+      
+      output <- optimal_cutpoint_object %>% 
+        summary
+      
+      output <- output$p.table$Global$AUC_CI
+      
+      output <- output %>% 
+        stri_replace(., fixed = ",", replacement = '') %>% 
+        stri_replace(., fixed = '(', replacement = '') %>% 
+        stri_replace(., fixed = ')', replacement = '') %>%
+        stri_split(., fixed = ' ', simplify = T) %>% 
+        as.numeric %>% 
+        c
+      names(output) <- c('AUC', 'lowerCI', 'upperCI')
+      
+      return(output)
+      
+    }
+    
+    plot_roc <- function(inp_table, 
[…]

### 2021-12-31 paper/markdown/figures.Rmd
Add files via upload

--- 2021-12-27 paper/markdown/figures.Rmd
@@
   \usepackage{xcolor}
-  \usepackage{float} \floatplacement{figure}{H}
+  \usepackage{float} \floatplacement{figure}{H} \floatplacement{table}{H}
   \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}       \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
@@
 
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, dev = "cairo_pdf", ft_max_row = 6, tab.topcaption = FALSE, ft.align = "center")
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, dev = "cairo_pdf", ft_max_row = 6, tab.topcaption = TRUE, ft.align = "center")
 
@@
 
-```{r tab-study-variables, tab.cap = 'Variables available for the analysis. The first 10 table records are shown. The full table is available as a supplementary Excel file.'}
+```{r tab-baseline-cohort, tab.cap = 'Baseline characteristic of the cohort.'}
 
-paper_tables$study_vars[1:10, ] %>% 
+paper_tables$baseline_cohort %>% 
   flextable() %>%
-  width(1, width = 4, unit = 'cm') %>% 
-  width(2:3, width = 2.5, unit = 'cm') %>% 
-  width(4:5, width = 4, unit = 'cm')
+  width(1, width = 3, unit = 'cm') %>% 
+  width(2, width = 7, unit = 'cm') %>% 
+  footnote(i = 2:7, 
+           j = 1, 
+           value = as_paragraph('percentage of the cohort (n individuals)'), 
+           part = 'body', 
+           ref_symbols = '1') %>% 
+  footnote(i = 4, 
+           j = 1, 
+           value = as_paragraph('body mass index, kg/m\u00B2, overweight: > 25 kg/m\u00B2, obesity: > 30 kg/m\u00B2'), 
+           part = 'body', 
+           ref_symbols = '2') %>% 
+  footnote(i = 5, 
+           j = 1, 
+           value = as_paragraph('packages per year'), 
+           part = 'body', 
+           ref_symbols = '3') %>% 
+  theme_vanilla
 
@@
 
-```{r tab-baseline-cohort, tab.cap = 'Baseline characteristic of the cohort.'}
-
-paper_tables$baseline_cohort %>% 
-  flextable() %>%
-  width(1, width = 3, unit = 'cm') %>% 
-  width(2, width = 7, unit = 'cm')
-
-```
-
-\newpage
-
-```{r tab-baseline-severity, tab.cap = 'Baseline characteristic of the acute COVID-19 severity subsets.
-Statistical significance of the differences between the severity subsets was assessed with $\\chi^2$ test.'}
-
-paper_tables$baseline_severity %>% 
-  flextable() %>%
-  width(1, width = 1.8, unit = 'cm') %>% 
-  width(2:5, width = 3.9, unit = 'cm') %>% 
-  width(6, width = 2.5, unit = 'cm')
-
-```
-
-\newpage
-
-```{r tab-ct-cohort, tab.cap = 'Lung CT features in the entire cohort at the consecutive follow-up visits.
-Statistical significance of the differences between the visits was assessed with $\\chi^2$ test.'}
+```{r tab-ct-cohort, tab.cap = 'Lung CT features in the entire cohort at the consecutive follow-up visits.'}
 
@@
   width(2:5, width = 3.2, unit = 'cm') %>% 
-  width(6, width = 2.5, unit = 'cm')
-
-```
-
-\newpage
-
-```{r tab-ct-visits, tab.cap = 'Lung CT features in the acute COVID-19 severity subsets at the consecutive follow-up visits.
-Statistical significance of the differences between the visits was assessed with $\\chi^2$ test.'}
-
-paper_tables$ct_visits %>% 
-  flextable() %>%
-  width(1, width = 1.8, unit = 'cm') %>% 
-  width(2, width = 2.7, unit = 'cm') %>% 
-  width(3:6, width = 3.2, unit = 'cm') %>% 
-  width(7, width = 2.5, unit = 'cm')
-
-```
-
-\newpage
-
-```{r tab-ct-severity, tab.cap = 'Lung CT features in the acute COVID-19 severity subsets at the consecutive follow-up visits.
-Statistical significance of the differences between the severity subsets was assessed with $\\chi^2$ test.'}
-
-paper_tables$ct_severity %>% 
-  flextable() %>%
-  width(1, width = 1.8, unit = 'cm') %>% 
-  width(2, width = 2.7, unit = 'cm') %>% 
-  width(3:6, width = 3.2, unit = 'cm') %>% 
-  width(7, width = 2.5, unit = 'cm')
-
-```
-
-\newpage
-
-```{r tab-ctss, tab.cap = 'CT severity score at the consecutive visits in the entire cohort and the acute COVID-19 severity subsets.
-Statistical significance of the differences between the visits was assessed with Kruskal-Wallis test.'}
-
-paper_tables$ctss %>% 
-  flextable() %>%
-  width(1, width = 1.8, unit = 'cm') %>% 
-  width(2:5, width = 3.5, unit = 'cm') %>% 
-  width(6, width = 2.5, unit = 'cm')
+  width(6, width = 2.5, unit = 'cm') %>% 
+  footnote(i = 1, 
+           j = 2:5, 
+           value = as_paragraph('percentage of the complete cases at the time point (n individuals), FUP: follow-up'), 
+           part = 'header', 
+           ref_symbols = '1') %>% 
+  footnote(i = 1, 
+           j = 6, 
+           value = as_paragraph('\u03C7\u00B2 test for differences between the follow-ups, ns: not significant'), 
+           part = 'header', 
+           ref_symbols = '2') %>% 
+  footnote(i = 6, 
+           j = 1, 
+           value = as_paragraph('ground glass opacity'), 
+           part = 'body', 
+           ref_symbols = '3') %>% 
+  footnote(i = 11, 
+           j = 1, 
+           value = as_paragraph('acute respiratory distress syndrome'), 
+           part = 'body', 
+           ref_symbols = '4') %>% 
+  footnote(i = 12, 
+           j = 1, 
+           value = as_paragraph('organizing pneumonia'), 
+           part = 'body', 
+           ref_symbols = '5') %>% 
+  theme_vanilla
 
@@
 
-```{r fig-consort, fig.width = mm_inch(paper_figures$ct_kinetic$w), fig.height = mm_inch(paper_figures$ct_kinetic$h), dpi = 600, fig.cap = 'Study inclusion flow diagram.'}
+```{r fig-consort, fig.width = mm_inch(paper_figures$ct_kinetic$w), fig.height = mm_inch(paper_figures$ct_kinetic$h), dpi = 600, fig.cap = 'Study inclusion flow diagram (placeholder).'}
 
@@
 
-__Figure \@ref(fig:fig-consort). Study inclusion flow diagram.__
+__Figure \@ref(fig:fig-consort). Study inclusion flow diagram (placeholder).__
 
@@
 
-```{r fig-ct-kinetic, fig.width = mm_inch(paper_figures$ct_kinetic$w), fig.height = mm_inch(paper_figures$ct_kinetic$h), dpi = 600, fig.cap = 'Resolution kinetic of any and moderate-to-severe lung CT abnormalities.'}
+```{r fig-photos, fig.width = mm_inch(paper_figures$ct_kinetic$w), fig.height = mm_inch(paper_figures$ct_kinetic$h), dpi = 600, fig.cap = 'Representative CT images (placeholder).'}
+
+ggdraw()
+
+```
+
+__Figure \@ref(fig:fig-photos). Representative CT images (placeholder).__
+
+\newpage
+
+```{r fig-ct-kinetic, fig.width = mm_inch(paper_figures$ct_kinetic$w), fig.height = mm_inch(paper_figures$ct_kinetic$h), dpi = 600, fig.cap = 'Resolution kinetic of any lung CT abnormalities.'}
 
@@
 
-__Figure \@ref(fig:fig-ct-kinetic). Resolution kinetic of any and moderate-to-severe lung CT abnormalities.__
+__Figure \@ref(fig:fig-ct-kinetic). Resolution kinetic of lung CT abnormalities.__
 
-Risk of any lung CT abnormalities (__A__) and moderate-to-severe lung CT abnormalities (CTSS > 5, __B__) at the consecutive time points was modeled by mixed-effect logistic regression (random effect: individual, fixed effect: time post COVID-19) in the entire cohort and the acute COVID-19 severity subsets.
+Risk of any lung CT abnormalities at the consecutive time points was modeled by mixed-effect logistic regression (random effect: individual, fixed effect: time post COVID-19) in the entire cohort and the acute COVID-19 severity subsets.
 Significance of the fixed time term was determined by likelihood ratio test (LRT) versus the respective random effect-only null model.
-Plots display percents of individuals with CT abnormalities in the analyzed collectives.
+
+__(A)__ Percents of individuals with CT abnormalities in the analyzed collectives.
 $\chi$ statistic, degrees of freedom and p values of the LRT test are presented in the plot captions.
 Numbers of individuals with the complete set of consecutive CT scans are shown under the plots.
+
+__(B)__ Recovery between the consecutive follow-ups assessed by Kendall $\tau$ statistic.
 
@@
 The effect size was determined by Kendall W test, differences between particular time points were compared by Wilcoxon test.
-Plots display individual CTSS value trajectories as thin gray lines, thick colored lines represent medians, interquartile ranges are presented as colored regions.
-$\chi$ statistic, degrees of freedom and p values of the Friedman test and the Kendall W statistic with 95$%$ confidence intervals are presented in the plot captions.
+
+__(A)__ Plots display individual CTSS value trajectories as thin gray lines, thick colored lines represent medians, interquartile ranges are presented as colored regions.
+$\chi$ statistic, degrees of freedom and p values of the Friedman test and the Kendall W statistic with 95$\%$ confidence intervals are presented in the plot captions.
 Numbers of individuals with the complete set of consecutive CT scans are shown under the plots.
+
+__(B)__ Recovery between the consecutive follow-ups assessed by Kendall $\tau$ statistic.
 
@@
 
-```{r fig-pili-risk, fig.width = mm_inch(paper_figures$pili_modeling$w), fig.height = mm_inch(paper_figures$pili_modeling$h), dpi = 600, fig.cap = 'Risk of developing PILI at the one-year post-COVID-19 follow-up.'}
+```{r fig-pili-risk, fig.width = mm_inch(paper_figures$pili_modeling$w), fig.height = mm_inch(paper_figures$pili_modeling$h), dpi = 600, fig.cap = 'Risk of developing persistent CT abnormalities at the one-year post-COVID-19 follow-up.'}
 
@@
 
-__Figure \@ref(fig:fig-pili-risk). Risk of developing PILI at the one-year post-COVID-19 follow-up.__
+__Figure \@ref(fig:fig-pili-risk). Risk of developing persistent CT abnormalities at the one-year post-COVID-19 follow-up.__
 
-__(A)__ Risk factors of developing PILI at the one-year follow-up visit were identified by a series of logistic models.
+__(A)__ Risk factors of developing any lung CT abnormality at the one-year follow-up visit were identified by a series of logistic models.
 Odds ratio (OR) significance was determined by Wald Z test.
-OR with 95$%$ confidence intervals are presented in a Forest plot.
+OR with 95$\%$ confidence intervals are presented in a Forest plot.
 Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.
 
-__(B)__ Independent risk factors of PILI were identified by multi-parameter logistic modeling with backward elimination.
+__(B)__ Independent risk factors of lung CT abnormality were identified by multi-parameter logistic modeling with backward elimination.
 OR significance was determined by Wald Z test.
 Model prediction was verified by 20-fold cross-validation (CV) and receiver-operating characteristic (ROC).
-OR with 95$%$ confidence intervals are presented in a Forest plot.
-Area under the curve (AUC), sensitivity (Se) and specificity (Sp) for detection of PILI in CV are indicated in the plot caption.
+OR with 95$\%$ confidence intervals are presented in a Forest plot.
+Area under the curve (AUC), sensitivity (Se) and specificity (Sp) for detection of lung CT abnormality in CV are indicated in the plot caption.
 Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.
 
-BMI: body mass index, PKY: packages per year.
+Orange: positive correlation, blue: negative correlation, gray: not significant, BMI: body mass index, PKY: packages per year, ref.: reference.
 
@@
 Regression coefficient ($\beta$) significance was determined by Wald Z test.
-Exponent $\beta$ with 95$%$ confidence intervals are presented in a Forest plot.
+Exponent $\beta$ with 95$\%$ confidence intervals are presented in a Forest plot.
 Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.
@@
 Model prediction was verified by 20-fold cross-validation (CV).
-Exponent $\beta$ with 95$%$ confidence intervals are presented in a Forest plot.
+Exponent $\beta$ with 95$\%$ confidence intervals are presented in a Forest plot.
 Mean absolute error (MAE) for prediction of CTSS and model R^2^ (Rsq) in CV are indicated in the plot caption.
@@
 
-BMI: body mass index, PKY: packages per year.
+Orange: positive correlation, blue: negative correlation, gray: not significant, BMI: body mass index, PKY: packages per year, ref.: reference.

### 2021-12-31 paper/markdown/supplementary_material.Rmd
Add files via upload

---
title: "Radio CovILD, 1-year follow-up"
subtitle: "Supplementary Material"
author: "CovILD study team"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
always_allow_html: true
output: 
  pdf_document:
    latex_engine: xelatex
bibliography: cov_biblio.bib
csl: elife.csl
header-includes:
  \usepackage{longtable}
  \usepackage{tabu}
  \usepackage{caption}
  \usepackage{makecell}
  \usepackage{pdflscape}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage{wrapfig}
  \usepackage{multirow}
  \usepackage[normalem]{ulem}
  \usepackage{colortbl}
  \usepackage{xcolor}
  \usepackage{float} \floatplacement{figure}{H} \floatplacement{table}{H}
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}       \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---

```{r, setup, include = FALSE}

library(bookdown)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, dev = "cairo_pdf", ft_max_row = 6, tab.topcaption = TRUE, ft.align = "center")

set_flextable_defaults(font.size = 11, padding = 2)

use_df_printer()

```

\beginsupplement
\newpage

# Supplementary Tables

```{r tab-study-variables, tab.cap = 'Variables available for the analysis. The first 10 table records are shown. The full table is available as a supplementary Excel file.'}

suppl_tables$study_vars[1:10, ] %>% 
  flextable() %>% 
  width(1, width = 4, unit = 'cm') %>% 
  width(2:3, width = 2.5, unit = 'cm') %>% 
  width(4:5, width = 4, unit = 'cm')

```

\newpage

```{r tab-baseline-severity, tab.cap = 'Baseline characteristic of the acute COVID-19 severity subsets.'}

suppl_tables$baseline_severity %>% 
  flextable() %>%
  width(1, width = 1.8, unit = 'cm') %>% 
  width(2:5, width = 3.9, unit = 'cm') %>% 
  width(6, width = 2.5, unit = 'cm') %>% 
  footnote(i = 2:6, 
           j = 1, 
           value = as_paragraph('percentage of the acute COVID-19 severity subset (n individuals)'), 
           part = 'body', 
           ref_symbols = '1') %>% 
  footnote(i = 4, 
           j = 1, 
           value = as_paragraph('body mass index, kg/m\u00B2, overweight: > 25 kg/m\u00B2, obesity: > 30 kg/m\u00B2'), 
           part = 'body', 
           ref_symbols = '2') %>% 
  footnote(i = 5, 
           j = 1, 
           value = as_paragraph('packages per year'), 
           part = 'body', 
           ref_symbols = '3') %>% 
  footnote(i = 1, 
           j = 6, 
           value = as_paragraph('Kruskal-Wallis test for differences between the acute COVID-19 severity subsets, ns: not significant'), 
           part = 'body', 
           ref_symbols = '4') %>% 
  footnote(i = 2:6, 
           j = 6, 
           value = as_paragraph('\u03C7\u00B2 test for differences between the acute COVID-19 severity subsets, ns: not significant'), 
           part = 'body', 
           ref_symbols = '5') %>% 
  theme_vanilla

```

\newpage

```{r tab-ct-visits, tab.cap = 'Lung CT features in the acute COVID-19 severity subsets at the consecutive follow-up visits.'}

suppl_tables$ct_visits %>% 
  flextable() %>%
  width(1, width = 1.8, unit = 'cm') %>% 
  width(2, width = 2.7, unit = 'cm') %>% 
  width(3:6, width = 3.2, unit = 'cm') %>% 
  width(7, width = 2.5, unit = 'cm') %>% 
  merge_v(1) %>% 
  footnote(i = 1, 
           j = 3:6, 
           value = as_paragraph('percentage of the acute COVID-19 severity subset (n individuals), FUP: follow-up'), 
           part = 'header', 
           ref_symbols = '1') %>% 
  footnote(i = 1, 
           j = 7, 
           value = as_paragraph('\u03C7\u00B2 test for differences between the follow-ups, ns: not significant'), 
           part = 'header', 
           ref_symbols = '2') %>% 
  footnote(i = c(6, 21, 36, 51), 
           j = 2, 
           value = as_paragraph('ground glass opacity'), 
           part = 'body', 
           ref_symbols = '3') %>% 
  footnote(i = c(11, 26, 41, 56), 
           j = 2, 
           value = as_paragraph('acute respiratory distress syndrome'), 
           part = 'body', 
           ref_symbols = '4') %>% 
  footnote(i = c(12, 27, 42, 57), 
           j = 2, 
           value = as_paragraph('organizing pneumonia'), 
           part = 'body', 
           ref_symbols = '5') %>% 
  theme_vanilla

```

\newpage

```{r tab-ctss, tab.cap = 'CT severity score at the consecutive visits in the entire cohort and the acute COVID-19 severity subsets.'}

suppl_tables$ctss %>% 
  flextable() %>%
  width(1, width = 1.8, unit = 'cm') %>% 
  width(2:5, width = 3.5, unit = 'cm') %>% 
  width(6, width = 2.5, unit = 'cm') %>% 
  footnote(i = 1, 
           j = 2:5, 
           value = as_paragraph('FUP: follow-up'), 
           part = 'header', 
           ref_symbols = '1') %>% 
  footnote(i = 1, 
           j = 6, 
           value = as_paragraph('Kruskal-Wallis test for differences between the follow-ups, ns: not significant'), 
           part = 'header', 
           ref_symbols = '2') %>% 
  theme_vanilla

```

\newpage

```{r tab-opacity, tab.cap = 'Percentages of opacity and high opacity at the consecutive visits in the entire cohort and the acute COVID-19 severity subsets.'}

suppl_tables$opacity %>% 
  flextable() %>%
  width(1, width = 1.8, unit = 'cm') %>% 
  width(2, width = 1.8, unit = 'cm') %>% 
  width(3:6, width = 3.5, unit = 'cm') %>% 
  width(7, width = 2.5, unit = 'cm') %>% 
  merge_v(1) %>% 
  footnote(i = 1, 
           j = 3:6, 
           value = as_paragraph('FUP: follow-up'), 
           part = 'header', 
           ref_symbols = '1') %>% 
  footnote(i = 1, 
           j = 7, 
           value = as_paragraph('Kruskal-Wallis test for differences between the follow-ups, ns: not significant'), 
           part = 'header', 
           ref_symbols = '2') %>% 
  theme_vanilla

```

\newpage

# Supplementary Figures

```{r fig-ct-sev, fig.width = mm_inch(suppl_figures$ct_kinetic$w), fig.height = mm_inch(suppl_figures$ct_kinetic$h), dpi = 600, fig.cap = 'Resolution kinetic of moderate-to-severe lung CT abnormalities.'}

suppl_figures$ct_kinetic$plot

```

__Figure \@ref(fig:fig-ct-sev). Resolution kinetic of moderate-to-severe lung CT abnormalities.__

Risk of moderate-to-severe lung CT abnormalities (CT severity score > 5) at the consecutive time points was modeled by mixed-effect logistic regression (random effect: individual, fixed effect: time post COVID-19) in the entire cohort and the acute COVID-19 severity subsets. Significance of the fixed time term was determined by likelihood ratio test (LRT) versus the respective random effect-only null model.

__(A)__ Percents of individuals with moderate-to-severe CT abnormalities in the analyzed collectives. $\chi$ statistic, degrees of freedom and p values of the LRT test are presented in the plot captions. Numbers of individuals with the complete set of consecutive CT scans are shown under the plots.

__(B)__ Recovery between the consecutive follow-ups assessed by Kendall $\tau$ statistic.

\newpage

```{r fig-ct-opacity, fig.width = mm_inch(suppl_figures$ct_opacity$w), fig.height = mm_inch(suppl_figures$ct_opacity$h), dpi = 600, fig.cap = 'Resolution kinetic of lung CT opacity.'}

suppl_figures$ct_opacity$plot

```

__Figure \@ref(fig:fig-ct-opacity). Resolution kinetic of lung CT opacity.__

Risk of any lung CT opacity at the consecutive time points was modeled by mixed-effect logistic regression (random effect: individual, fixed effect: time post COVID-19) in the entire cohort and the acute COVID-19 severity subsets. Significance of the fixed time term was determined by likelihood ratio test (LRT) versus the respective random effect-only null model.

__(A)__ Percents of individuals with CT opacity in the analyzed collectives. $\chi$ statistic, degrees of freedom and p values of the LRT test are presented in the plot captions. Numbers of individuals with the complete set of consecutive CT scans are shown under the plots.

__(B)__ Recovery between the consecutive follow-ups assessed by Kendall $\tau$ statistic.

\newpage

```{r fig-opacity-kinetic, fig.width = mm_inch(suppl_figures$opacity_kinetic$w), fig.height = mm_inch(suppl_figures$opacity_kinetic$h), dpi = 600, fig.cap = 'Kinetic of the lung opacity.'}

suppl_figures$opacity_kinetic$plot

```

__Figure \@ref(fig:fig-opacity-kinetic). Kinetic of the lung opacity.__

Kinetic of the lung opacity at the consecutive time points was investigated by Friedman test (grouping by the individual) in the entire cohort and the acute COVID-19 severity subsets. The effect size was determined by Kendall W test, differences between particular time points were compared by Wilcoxon test.

__(A)__ Plots display individual opacity trajectories as thin gray lines, thick colored lines represent medians, interquartile ranges are presented as colored regions. $\chi$ statistic, degrees of freedom and p values of the Friedman test and the Kendall W statistic with 95$\%$ confidence intervals are presented in the plot captions. Numbers of individuals with the complete set of consecutive CT scans are shown under the plots.

__(B)__ Recovery between the consecutive follow-ups assessed by Kendall $\tau$ statistic.

\newpage

```{r fig-opacity-risk, fig.width = mm_inch(suppl_figures$opacity_risk$w), fig.height = mm_inch(suppl_figures$opacity_risk$h), dpi = 600, fig.cap = 'Risk of developing lung opacity at the one-year post-COVID-19 follow-up.'}

suppl_figures$opacity_risk$plot

```

__Figure \@ref(fig:fig-opacity-risk). Risk of developing lung opacity at the one-year post-COVID-19 follow-up.__

__(A)__ Risk factors of developing opacity at the one-year follow-up visit were identified by a series of logistic models. Odds ratio (OR) significance was determined by Wald Z test. OR with 95$\%$ confidence intervals are presented in a Forest plot. Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.

__(B)__ Independent risk factors of opacity were identified by multi-parameter logistic modeling with backward elimination. OR significance was determined by Wald Z test. Model prediction was verified by 20-fold cross-validation (CV) and receiver-operating characteristic (ROC). OR with 95$\%$ confidence intervals are presented in a Forest plot. Area under the curve (AUC), sensitivity (Se) and specificity (Sp) for detection of opacity in CV are indicated in the plot caption. Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.

Orange: positive correlation, blue: negative correlation, gray: not significant, BMI: body mass index, PKY: packages per year, ref.: reference.

\newpage

```{r fig-opacity-rater, fig.width = mm_inch(suppl_figures$any_abnormality_rater$w), fig.height = mm_inch(suppl_figures$any_abnormality_rater$h), dpi = 600, fig.cap = 'Performance of automated opacity assessment at detection of lung CT abnormalities.'}

suppl_figures$any_abnormality_rater$plot

```

__Figure \@ref(fig:fig-opacity-rater). Performance of automated opacity assessment at detection of lung CT abnormalities.__

Lung CT abnormality were defined either by the radiologist or by the software as opacity $\geq$ 1$\%$ lung volume. Performance of the opacity at detecting CT abnormalities was investigated at the consecutive follow-ups by receiver-operating characteristic (ROC) and Cohen's $\kappa$ statistic. Numbers of complete observations and CT abnormalities (events) are indicated under the plots.

__(A)__ ROC curves annotated with sensitivity (Se), specificity (Sp) and area under the curve (AUC) values with 95$\%$ confidence intervals.

__(B)__ $\kappa$ with 95$\%$ confidence intervals presented in a forest plot.

\newpage

```{r fig-ggo-rater, fig.width = mm_inch(suppl_figures$ggo_rater$w), fig.height = mm_inch(suppl_figures$ggo_rater$h), dpi = 600, fig.cap = 'Performance of automated opacity assessment at detection of GGO.'}

suppl_figures$ggo_rater$plot

```

__Figure \@ref(fig:fig-ggo-rater). Performance of automated opacity assessment at detection of GGO.__

Lung GGO were defined either by the radiologist or by the software as opacity $\geq$ 1$\%$ lung volume. Performance of the opacity at detecting GGO was investigated at the consecutive follow-ups by receiver-operating characteristic (ROC) and Cohen's $\kappa$ statistic. Numbers of complete observations and CT abnormalities (events) are indicated under the plots.

__(A)__ ROC curves annotated with sensitivity (Se), specificity (Sp) and area under the curve (AUC) values with 95$\%$ confidence intervals.

__(B)__ $\kappa$ with 95$\%$ confidence intervals presented in a forest plot.

\newpage

```{r fig-ctss-opacity, fig.width = mm_inch(suppl_figures$ctss_opacity$w), fig.height = mm_inch(suppl_figures$ctss_opacity$h), dpi = 600, fig.cap = 'Correlation of CTSS and lung opacity.'}

suppl_figures$ctss_opacity$plot

```

__Figure \@ref(fig:fig-ctss-opacity). Correlation of CTSS and lung opacity.__

Correlation of CTSS and lung opacity at the consecutive follow-ups was investigated with Spearman test. Points represent single observations, blue lines with gray regions represent fitted linear trend with 95$\%$ confidence interval. Correlation coefficients (r), p values and numbers of complete observations are indicated in the plot captions.

### 2021-12-31 analysis scripts/interrater.R

@@ -5,157 +5,23 @@
 # container list ----
   
   rater <- list()
-
-# inter-rater comparison, any abnormalities by CTSS and opacity ----
   
-  insert_msg('Any abnormalities')
-
-  ## table objects
-    
-  rater$any_ct$tables <- radio$long %>% 
-    rater_tbl(ctss_any, opacity) %>% 
-    map(~table(.x[c('ctss_any', 'opacity')]))
+# inter-rater comparison, any abnormalities, GGO, consolidations by CTSS and opacity ----
   
-  ## n numbers
+  insert_msg('Any abnormalities and opacity')
   
-  rater$any_ct$n_numbers <- map2(map(rater$any_ct$tables, sum), 
-                                 map(rater$any_ct$tables, ~sum(.x[2, ])), 
-                                 ~paste0('total: n = ', .x, ', events: n = ', .y))
-  
-  rater$any_ct$n_tag <- map2_chr(globals$visit_labels, 
-                                 rater$any_ct$n_numbers, 
-                                 paste, sep = '\n')
-  
-  ## Kappa calculation
-  
-  rater$any_ct$kappa <- rater$any_ct$tables %>% 
-    map(Kappa_pipe) %>% 
-    map2_dfr(., names(.), ~mutate(.x, visit = .y)) %>% 
-    mutate(visit = factor(visit, levels = rev(visit)))
-  
-  ## ROC
-  
-  rater$any_ct$cutpoint <- radio$long %>% 
-    rater_tbl(ctss_any, opacity) %>% 
-    map(~map_dfc(.x, ~as.numeric(.x) - 1)) %>% 
-    map(find_optimal_cutoff, 
-        status_variable = 'ctss_any', 
-        marker = 'opacity')
-  
-  rater$any_ct$roc_stats <- rater$any_ct$cutpoint %>% 
-    map(get_optimal_cutpoint_coefs) %>% 
-    map(unlist, recursive = FALSE) %>% 
-    map(~.x[2:3]) %>%
-    as_tibble %>% 
-    t %>% 
-    as.data.frame %>% 
-    set_names(c('Se', 'Sp')) %>% 
-    rownames_to_column('visit')
-  
-  rater$any_ct$roc_stats <- rater$any_ct$cutpoint %>% 
-    map_dfr(get_auc) %>% 
-    cbind(rater$any_ct$roc_stats, .) %>% 
-    as_tibble %>% 
-    mutate(plot_annotation = paste0('Se = ', signif(Se, 2), 
[…]

### 2021-12-31 analysis scripts/kinetic.R

@@ -46,27 +46,27 @@
   plan('multisession')
   
   kinetic$numeric$effsize <- list(ctss = ctss ~ visit|ID, 
-                               perc_opac = perc_opac ~ visit|ID, 
-                               perc_hiopac = perc_hiopac ~ visit|ID) %>% 
+                                  perc_opac = perc_opac ~ visit|ID, 
+                                  perc_hiopac = perc_hiopac ~ visit|ID) %>% 
     map(function(outcome) kinetic$analysis_tbl %>% 
           future_map(friedman_effsize, 
                      formula = outcome, 
                      ci = TRUE, 
-                     ci.type = 'bca', 
+                     ci.type = 'norm', 
                      .options = furrr_options(seed = TRUE)) %>% 
           map2_dfr(., names(.), ~mutate(.x, subset = .y)))
-
+  
   kinetic$numeric$summary <- map2(kinetic$numeric$summary, 
                                   kinetic$numeric$effsize, 
                                   ~mutate(.x, 
                                           w = paste0(signif(.y$effsize, 2), 
                                                      ' [', signif(.y$conf.low, 2), 
                                                      ' - ', signif(.y$conf.high, 2), ']'), 
-                                          plot_caption = paste(plot_caption, w, sep = ', W = ')))
+                                          plot_caption = paste(plot_caption, w, sep = '\nW = ')))
 
   plan('sequential')
   
-  ## post-hoc tests
+  ## post-hoc Wilcoxon tests
   
   kinetic$numeric$post_hoc <- list(ctss = ctss ~ visit, 
                                    perc_opac = perc_opac ~ visit, 
@@ -78,7 +78,19 @@
                      method = 'wilcox.test', 
                      p.adjust.method = 'holm'))
   
-  ## plots
+  ## post-hoc effect size
+  
+  kinetic$numeric$post_hoc_eff_size <- c('ctss', 'perc_opac', 'perc_hiopac') %>% 
+    map(function(outcome) kinetic$analysis_tbl %>% 
+          map(compare_effects, 
+              response = outcome, 
+              indep_var = 'visit', 
+              group_var = 'ID', 
+              invert = TRUE) %>% 
+          map2_dfr(., names(.), function(data, sub) mutate(data, severity = sub))) %>% 
+    set_names(c('ctss', 'perc_opac', 'perc_hiopac'))
+  
[…]

### 2021-12-31 paper scripts/deploy_paper.R

@@ -14,36 +14,6 @@
   
 # Specific functions -----
   
-  tab_preview <- function(inp_tbl, head_lines = 'all', signif_digits = 3) {
-    
-    ## makes a nice table head for preview in the figure file
-    
-    out_tbl <- inp_tbl %>% 
-      map_dfc(function(x) if(is.numeric(x)) signif(x, digits = signif_digits) else x)
-    
-    if(head_lines != 'all') {
-      
-      return(out_tbl[1:head_lines, ])
-      
-    } else {
-      
-      return(out_tbl)
-      
-    }
-    
-  }
-  
-  collapse_and <- function(vector, collapse = ', ', last_link = ' and ') {
-    
-    ## collapses a given vector with the last_link string before the last item
-    
-    coll_phrase <- paste(vector[1:length(vector) - 1], collapse = collapse) %>% 
-      paste(., vector[length(vector)], sep = last_link)
-    
-    return(coll_phrase)
-    
-  }
-  
   mm_inch <- function(input_mm) {
     
     return(0.0393700787 * input_mm)
@@ -55,6 +25,10 @@
   insert_msg('Rendering figure and table legend file')
   
   render('./paper/markdown/figures.Rmd', 
+         output_format = word_document2(number_sections = F, theme = 'readable'), 
+         output_dir = './paper/')
+  
+  render('./paper/markdown/supplementary_material.Rmd', 
          output_format = word_document2(number_sections = F, theme = 'readable'), 
          output_dir = './paper/')
 

### 2021-12-31 paper scripts/paper_figures.R

@@ -11,9 +11,13 @@
   
   insert_msg('Figure 1: CONSORT, provided by the authors')
   
-# Figure 2: kinetics of any CT abnormalities -----
+# Figure 2: CT images, provided by the Author team -----
   
-  insert_msg('Figure 2: kinetic of any CT abnormalities and CTSS')
+  insert_msg('Figure 2: CT images, provided by the authors')
+  
+# Figure 3: kinetics of any CT abnormalities -----
+  
+  insert_msg('Figure 3: kinetic of any CT abnormalities and CTSS')
   
   paper_figures$ct_kinetic$upper_panel <- kinetic$binary$plots[c('cohort.ctss_any', 
                                                                  'mild.ctss_any', 
@@ -29,52 +33,56 @@
                               breaks = c(2, 3, 6, 12))) %>% 
     plot_grid(plotlist = ., 
               ncol = 3, 
-              align = 'hv')
+              align = 'hv', 
+              axis = 'tblr')
   
-  paper_figures$ct_kinetic$bottom_panel <- kinetic$binary$plots[c('cohort.ctss_mod_severe', 
-                                                                   'mild.ctss_mod_severe', 
-                                                                   'moderate.ctss_mod_severe', 
-                                                                   'severe.ctss_mod_severe', 
-                                                                   'critical.ctss_mod_severe')] %>% 
-    map2(., globals$subset_labels, 
-         ~.x + 
-           labs(title = .y) + 
-           scale_y_continuous(limits = c(0, 115), 
-                              breaks = seq(0, 100, by = 25)) + 
-           scale_x_continuous(limits = c(2, 13), 
-                              breaks = c(2, 3, 6, 12))) %>% 
-    plot_grid(plotlist = ., 
-              ncol = 3, 
-              align = 'hv')
+  paper_figures$ct_kinetic$bottom_panel <- kinetic$binary$eff_plots$ctss_any + 
+    theme(legend.position = 'none') + 
+    facet_grid(.~severity, 
+               labeller = as_labeller(globals$subset_labels))
+
   
   paper_figures$ct_kinetic <- plot_grid(paper_figures$ct_kinetic$upper_panel, 
                                         paper_figures$ct_kinetic$bottom_panel, 
                                         nrow = 2, 
+                                        rel_heights = c(2, 1), 
                                         labels = LETTERS, 
                                         label_size = 10) %>% 
-    as_figure_object(figure_label = 'figure_2_kinetic_ct', 
[…]

### 2021-12-31 paper scripts/paper_tables.R

@@ -5,126 +5,109 @@
 # container list ----
  
  paper_tables <- list()
+ suppl_tables <- list()
  
-# Table 1: Study variables -----
+# Table 1: baseline cohort characteristic -----
  
- insert_msg('Table 1: study variables')
- 
- paper_tables$study_vars <- radio$vars %>% 
-   select(old_var, 
-          new_var, 
-          label, 
-          description, 
-          modeling) %>% 
-   mutate(modeling = car::recode(modeling, "'no' = 'not used in modeling'; 'long_outcome' = 'outcome'")) %>% 
-   set_names(c('SPSS variable name', 
-               'R variable name', 
-               'R variable label', 
-               'Description', 
-               'Variable type'))
- 
-# Table 2: baseline cohort characteristic -----
- 
- insert_msg('Table 2: baseline cohort characteristic')
+ insert_msg('Table 1: baseline cohort characteristic')
  
  paper_tables$baseline_cohort <- eda$baseline$summary_cohort %>% 
-   select(label, statistic) %>% 
-   set_names(c('Variable', 'Statistic'))
+    select(label, statistic) %>% 
+    map_dfc(stri_replace, regex = '\\ncomplete.*', replacement = '') %>% 
+    set_names(c('Variable', paste('Statistic, n =', nrow(radio$clear))))
  
-# Table 3: baseline characteristic, severity groups -----
+# Table 2: CT abnormalities at the visits, whole cohort ----
  
- insert_msg('Table 3: baseline characteristic, severity groups')
- 
- paper_tables$baseline_severity <- eda$baseline$summary_severity %>% 
-   select(label, 
-          mild, 
-          moderate, 
-          severe, 
-          critical, 
-          significance) %>% 
-  set_names(c('Variable', 
-              'Mild', 
-              'Moderate',
-              'Severe', 
-              'Critical', 
-              'Significance'))
- 
-# Table 4: CT abnormalities at the visits, whole cohort ----
- 
- insert_msg('Table 4: CT abnormalities, whole cohort, consecutive visits')
+ insert_msg('Table 2: CT abnormalities, whole cohort, consecutive visits')
  
  paper_tables$ct_cohort <- eda$ct$summary_cohort %>% 
-   select(label, 
-          starts_with('fup'), 
-          significance) %>% 
-   set_names(c('Variable', 
-               '2-month FUP', 
-               '3-month FUP', 
-               '6-month FUP', 
-               '1-year FUP', 
[…]

### 2021-12-31 tools/project_globals.R

@@ -40,11 +40,11 @@
   
   insert_msg('Variable representation')
 
-  globals$subset_colors <- c('cohort' = 'gray30', 
+  globals$subset_colors <- c('cohort' = 'plum3', 
                              'mild' = 'steelblue', 
                              'moderate' = 'cornsilk4', 
                              'severe' = 'coral3', 
-                             'critical' = 'firebrick')  
+                             'critical' = 'firebrick3')  
   
   globals$subset_labels <- c(cohort = 'Cohort', 
                              mild = 'Mild COVID-19', 

### 2021-12-31 tools/project_tools.R

@@ -6,6 +6,7 @@
   require(stringi)
   require(foreign)
   require(vcd)
+  require(rcompanion)
   
 # data import and transformation -----
   
@@ -254,7 +255,248 @@
     
   }
   
-# kinetic plotting -----
+# kinetic modeling and plotting -----
+  
+  compare_effects <- function(data, 
+                              response = 'ctss', 
+                              indep_var = 'visit', 
+                              group_var = 'ID',
+                              invert = FALSE, 
+                              parallel = FALSE, 
+                              ci.type = 'norm') {
+    
+    ## pairwise effect size
+    
+    start_time <- Sys.time()
+    on.exit(message(paste('Elapsed:', Sys.time() - start_time)))
+    
+    pairs <- combn(unique(data[[indep_var]]), m = 2, simplify = FALSE)
+    
+    group_tbl <- pairs %>% 
+      map_dfr(~tibble(.y = response, 
+                      group1 = .x[1], 
+                      group2 = .x[2]))
+    
+    mod_tbl <- pairs %>% 
+      map(~filter(data, .data[[indep_var]] %in% .x)) %>% 
+      map(arrange, .data[[group_var]]) %>% 
+      map(mutate, 
+          x_fct = droplevels(.data[[indep_var]]))
+    
+    N <- mod_tbl %>% 
+      map(filter, !duplicated(.data[[group_var]])) %>% 
+      map_dbl(nrow)
+    
+    ## pairwise effect size calculation
+  
+    my_kendall <- function(x, y) {
+      
+      kendall_res <- try(KendallTauB(x, y, conf.level = 0.95), silent = TRUE)
+      
+      if(any(class(kendall_res) == 'try-error')) {
+        
+        warning('The factor value is invariant!', call. = FALSE)
+        
+        kendall_res <- c(tau_b = 0, lwr.ci = NA, upr.ci = NA)
+        
+      }
+      
+      return(kendall_res)
+      
+    }
+    
+    my_freeman <- function(tbl) {
+      
+      freeman <- try(freemanTheta(tbl, 
+                                  ci = TRUE, 
+                                  type = ci.type, 
+                                  R = 1000), 
+                     silent = TRUE)
+      
+      if(any(class(freeman) == 'try-error')) {
+        
+        warning('The factor value is invariant!', call. = FALSE)
+        
+        freeman <- tibble(Freeman.theta = 0, 
+                          lower.ci = NA, 
+                          upper.ci = NA)  
+      }
+      
+      return(freeman)
+      
+    }
+    
[…]

### 2022-02-08 tools/counting_tools.R
Not needed any more, replaced by a package

@@ -1,1077 +1,0 @@
-# This simple script contains some tools for counting variable levels
-# And calculating the characteristic of numeric variables
-
-
-# libraries -----
-
-  require(plyr)
-  require(tidyverse)
-  require(stringi)
-  require(ggsignif)
-  require(ggpubr)
-  
-# functions for factors -------
-
-  count_feature <- function(inp_tbl, var_to_count, remove_na = T, .drop = T) {
-    
-    ## calculates the percentage and number of participants with/without the given feature
-    ## the .drop argument specifies if the empty factor levels will be included in the output
-    
-    if(remove_na) {
-      
-      count_tbl <- inp_tbl %>% 
-        filter(!is.na(.data[[var_to_count]]), 
-               .data[[var_to_count]] != 'no_answer')
-      
-    } else {
-      
-      count_tbl <- inp_tbl
-      
-    }
-    
-    feature_counts <- count_tbl %>% 
-      count(.data[[var_to_count]], 
-            .drop = .drop) %>% 
-      mutate(percent = n/sum(n) * 100, 
-             total_n = sum(n))
-    
-    return(feature_counts)
-    
-  }
-
-  count_feature_lst <- function(inp_tbl, var_to_count_vec, remove_na = T, positive_only = T) {
-    
-    ## wrapper for variable vectors
-    
-    count_tbl <- var_to_count_vec %>% 
-      map(count_feature, 
-          inp_tbl = inp_tbl, 
-          remove_na = remove_na) %>% 
-      map(set_names, 
-          c('feature_strata', 
-            'n', 
-            'percent', 
-            'total_n')) %>% 
-      set_names(var_to_count_vec) %>% 
-      map2_dfr(., names(.), 
-               function(x, y) mutate(x, feature = y))
-    
-    if(positive_only) {
-      
-      count_tbl <- count_tbl %>% 
-        filter(feature_strata == 'yes')
-      
-    }
-    
-    return(count_tbl)
-    
-  }
-  
-  pie_feature <- function(perc_tbl, plot_title = names(perc_tbl)[1], plot_subtitle = NULL, 
-                          order_tbl = NULL, pie = T, repel = T, fill_colors = NULL, cust_theme = NULL) {
-    
-    ## plots the distribution of the given feature: takes an output of the count_feature function
-    ## a tibble with order for plotting may be provided (names: plot_order, feature_strata)
-    
-    plotting_tbl <- perc_tbl %>% 
-      set_names(c('feature_strata', 
[…]

### 2022-02-08 tools/lm_qc_tools.R
Not needed any more, replaced by a package

@@ -1,483 +1,0 @@
-# This script provides tools for summaries and QC of linear (lm) and generalized linear models
-
-# libraries ------
-
-  require(tidyverse)
-  require(ggrepel)
-  require(broom)
-  require(furrr)
-  require(stringi)
-
-# auxiliary functions ----
-
-  point_plot_ <- function(data, x_var, y_var, x_lab = x_var, y_lab = y_var, 
-                          plot_title = NULL, smooth = T, silent = T, ...) {
-    
-    ## draws a simple point plot for diagnostic purposes, takes the output of get_qc_tbl() as data argument
-    ## color-codes model missfits
-    
-    ## table for plotting 
-    
-    data <- data %>% 
-      mutate(misslab = try(ifelse(.candidate_missfit == 'yes',
-                                  .rownames, 
-                                  NA), silent = T),
-             misslab = ifelse(any(class(misslab) == 'try-error'), NA, misslab))
-    
-    ## fill colors
-    
-    fill_colors <- c(no = 'cornflowerblue', 
-                     yes = 'firebrick4')
-    
-    ## point plot
-    
-    point_plot <- data %>% 
-      ggplot(aes(x = .data[[x_var]], 
-                 y = .data[[y_var]], 
-                 fill = .candidate_missfit)) + 
-      geom_point(size = 2, 
-                 shape = 21) + 
-      geom_text_repel(aes(label = misslab), 
-                      show.legend = F) + 
-      scale_fill_manual(values = fill_colors, 
-                        name = 'Candidate missfit') + 
-      labs(x = x_lab, 
-           y = y_lab, 
-           title = plot_title)
-    
-    if(smooth) {
-      
-      if(silent) {
-        
-        suppressWarnings(point_plot <- point_plot + 
-                           geom_smooth(show.legend = F, 
-                                       color = 'black', 
-                                       fill = 'dodgerblue2', ...))
-        
-      } else {
-        
-        point_plot <- point_plot + 
-          geom_smooth(show.legend = F, 
-                      color = 'black', 
-                      fill = 'dodgerblue2', ...)
-        
-      }
-      
-    }
-      
-    return(point_plot)
-    
-  }
-  
-  calc_expected_ <- function(inp_data, observed) {
-    
-    ## calculates expected normal distribution of a variable observed
[…]

### 2022-02-08 tools/modeling_tools.R
Delete modeling_tools.R

@@ -1,526 +1,0 @@
-# This script provides functional tools for modeling tasks in christas Sepsis project
-
-# libraries -----
-
-  library(plyr)
-  library(dplyr)
-  library(tidyr)
-  library(purrr)
-  library(lme4)
-  library(lmerTest)
-  library(stringi)
-
-# globals ----
-
-  default_signif = 2
-
-# simple model generation functions -----
-
-  make_lmer <- function(inp_table, response = NULL, fix_variables = NULL, random_variables = NULL, interaction_term = '', ...) {
-    
-    # generaes a simple no-interaction lm (if random variables not given) or lmer
-    # Interaction can be included by explicitly stating the interaction term
-    
-    inp_formula <- paste(response, '~', paste(fix_variables, collapse = '+'), interaction_term, sep = '+')
-    
-    if(is.null(random_variables)) {
-      
-      out_model <- lm(formula = as.formula(inp_formula), data = inp_table, ...)
-      
-    } else {
-      
-      for(var in random_variables) {
-        
-        inp_formula <- paste(inp_formula, '+ (1|', var, ')')
-        
-      }
-      
-      out_model <- lmer(formula = as.formula(inp_formula), data = inp_table, ...)
-      
-    }
-    
-    return(out_model)
-    
-  }
-  
-  make_glm <- function(inp_table, response = NULL, fix_variables = NULL, ...) {
-    
-    # generaes a simple no-interaction glm
-    
-    inp_formula <- paste(response, '~', paste(fix_variables, collapse = '+'))
-    
-    return(glm(formula = as.formula(inp_formula), data = inp_table, ...))
-    
-  }
-  
-# model summary functions ----
-
-  get_lme_results <- function(model_mix, L_matrix = NULL, anova = F, ...){
-    
-    require(lmerTest)
-    
-    ## a function returning confidence intervals and significances for particular coefficents of the mixed model with one fixed
-    ## effect. The coefficients to display are defined by the L_matrix. Uses the contest function from the lmerTest package
-    ## If the L_matrix is not given, stats for all coefficients are calculated
-    
-    output <- model_mix %>% 
-      as_lmerModLmerTest
-    
-    if(is.null(L_matrix)) {
-      
-      L_matrix <- identity_matrix(length(model_mix@beta))
-      
-    }
-    
-    if(anova) {
-      
-      output <- contest(output, L_matrix, ...)
-      
-    } else {
-      
-      output <- contest(output, L_matrix, joint = F, ...) %>% 
[…]

### 2022-02-08 analysis scripts/risk_score.R
Modeling strategy changed
Requested by the statistical reviewer

@@ -1,211 +1,0 @@
-# This script models risk of any, moderate-severe, severe, ggo, reticulation, consolidations and bronchial dilatation 
-# (logistic regression) as well as of the abnormality grading at the 1-year follow-up. Explanatory variables
-# are> sex, age, severity, bin, smoking parameters as well as initial CT abnormality severity (CTSS with the cutoffs 5 and 10)
-# Multi-parameter models are constructed by backward elimination and validated by 20-fold CV
-
-  insert_head()
-  
-# container list -----
-  
-  risk <- list()
-
-
-# globals ------
-  
-  insert_msg('Globals')
-  
-  risk$explanatory_vars <- c(radio$long_expl_var)
-  
-  risk$outcomes <-  c('ggo_fup4', 
-                      'retic_fup4', 
-                      'ctss_any_fup4', 
-                      'opacity_fup4', 
-                      'ctss_fup4', 
-                      'log_perc_opac_fup4')
-  
-  risk$outcome_forms <- risk$outcomes %>% 
-    map(~paste(.x, paste(risk$explanatory_vars, collapse = '+'), sep = '~')) %>% 
-    map(as.formula)
-  
-  risk$families <- c(rep('binomial', 4), 
-                     'poisson', 
-                     'gaussian')
-  
-  risk$analysis_tbl <- radio$clear %>% 
-    mutate(log_perc_opac_fup4 = log(perc_opac_fup4 + 1))
-  
-  for(i in risk$outcomes) {
-    
-    if(is.factor(risk$analysis_tbl[[i]])) {
-      
-      risk$analysis_tbl <- risk$analysis_tbl %>% 
-        mutate(!!sym(i) := as.numeric(.data[[i]]) - 1)
-      
-    }
-    
-  }
-  
-  risk$levene_forms <- risk$explanatory_vars %>% 
-    paste('.resid ~', .) %>% 
-    map(as.formula)
-  
-# serial univariate modeling -----
-
-  insert_msg('Serial univariate modeling')  
-  
-  risk$univariate_models <- list(response = risk$outcomes, 
-                                 family = risk$families) %>% 
-    pmap(make_lm_model, 
-        data = risk$analysis_tbl, 
-        indep_variable = risk$explanatory_vars, 
-        mod_fun = glm, 
-        est_transf = exp, 
-        error_resistant = FALSE) %>% 
-    set_names(risk$outcomes)
-
[…]

### 2022-02-08 paper/markdown/response_stat_reviewer.Rmd
Paper revision
Modeling strategy changes, figure appearance modified

---
title: "Radio CovILD, 1-year follow-up"
subtitle: "Response to the statistical reviewer"
author: "CovILD study team"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
always_allow_html: true
output: 
  pdf_document:
    latex_engine: xelatex
bibliography: cov_biblio.bib
csl: elife.csl
header-includes:
  \usepackage{longtable}
  \usepackage{tabu}
  \usepackage{caption}
  \usepackage{makecell}
  \usepackage{pdflscape}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage{wrapfig}
  \usepackage{multirow}
  \usepackage[normalem]{ulem}
  \usepackage{colortbl}
  \usepackage{xcolor}
  \usepackage{float} \floatplacement{figure}{T} \floatplacement{table}{H}
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{R\arabic{table}}       \setcounter{figure}{0} \renewcommand{\thefigure}{R\arabic{figure}}}
---

```{r, setup, include = FALSE}

library(bookdown)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, dev = "cairo_pdf", ft_max_row = 6, tab.topcaption = TRUE, ft.align = "center")

set_flextable_defaults(font.size = 11, padding = 2)

use_df_printer()

```

\beginsupplement

\newpage

# Response to the statistical reviewer

We would like to thank the reviewer for careful study of the manuscript and the feedback. Please find below our response to the open issues.

## Issue 1

__Issue:__ In general, the statistical methods and results need pruning. There are too many analyses for this single paper, and aren’t needed for the main aims seem to be focused around the inferential analysis of the risk factors associated with any CT abnormalities (yes/no) and the risk factors associated with CT severity score (0-25). Also, I see the benefit of understanding the trend of CTSS over time.

__Response:__ We reduced the amount of data and analyses presented in the manuscript as suggested and followed the specific comments of the editor in the manuscript text. In particular, __Figure 6__ comping with the inter-rater comparison of the manual and automated abnormality detection and the plots of effect sizes in __Figure 5B__ are removed and the effect size measures cited in the main text instead.

## Issue 2

__Issue:__  Methods and results related to the automatic quantification opacity detection seems outside of the scope of this paper. Removing the text and results relating to this would really help the readability of this paper, and make the results clearer.

__Response__: The corresponding text content and __Figure 6__ were removed from the revised manuscript.

```{r fig-age-spline, fig.width = 7.086614166, fig.height = 8.267716527, fig.cap = 'Effects of the age variable splining.'}

rev_figures$age_spline$plot

```

__Figure R\@ref(fig:fig-age-spline). Effects of the age variable splining.__

## Issue 3

__Issue:__ Instead of categorization of the continuous variables, you should considering splining these variables in the regression analyses. Even if the variables are not normal there are better and more powerful ways of handling these variables in the analysis.

__Response:__ This is certainly an important point. We agree that manual categorization of numeric variables is controversial and not optimal in terms of model performance. However, categorization of clinical and demographic parameters with widely accepted/published cutoffs is usually accepted better by the clinical community. In our manuscript, we applied cutoffs of BMI and pack-years associated with clinical phenotypes (see: [@Pleasants2020] for smoking). For the age variable, the cutoff choice was motivated by clinical experience of the study team and corresponded approvimately to the mean age of the study cohort.

To address the issue in more detail, we modeled the risk of any CT abnormalities as a function of splined age (logistic GAM, thin plate smoothing method, k = 20, R package _mgcv_, function _gam()_). Interestingly, the inflection point of the predicted risk curve approximates well the chosen age cutoff of 60-years (__Figure R\@ref(fig:fig-age-spline)A__). In general, inclusion of categorized age or smoothed age term in the multi-variate risk modeling (explanatory variables: age, sex and COVID-19 severity) has negligible effect on the accuracy, sensitivity and specificity at predicting any CT abnormalities at the 1-year follow-up (__Figure R\@ref(fig:fig-age-spline)B__).

The comparison of the smoothed and cutoff models are included in the project pipeline deposited at GitHub (https://github.com/PiotrTymoszuk/radio_CovILD).

## Issue 4

__Issue:__ The univariable models are fine on their own (and I think the associated plots are super nice!), but they should not be used to inform the variables that are included in your multivariable models.

__Response:__ We apologize for the unclarity. The multi-parameter models were constructed with the full set of candidate variables (i.e. age, sex, BMI class, smoking quantity, smoking status and acute COVID-19 severity) without any pre-selection based on the results of univariable modeling. The sole criterion for full model term elimination was AIC reduction  (function _stepAIC()_, R package _MASS_). We clarify this in the _Statistical analysis_ section.

```{r fig-pois-ord-comp, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Comparison of the Poisson and logistic ordinal regression results.'}

rev_figures$pois_ord_comp$plot

```

__Figure R\@ref(fig:fig-pois-ord-comp). Comparison of the Poisson and logistic ordinal regression results.__

## Issue 5

__Issue:__ Poisson regression doesn’t seem quite right for the CTSS outcome. Ordinal logistic regression would be better since certain outcomes are "higher" or "lower" than others, but you're not sure by how much or if it even matters how much. In this example, the outcomes are ratings on a 1-25 scale, you know 4 is better than 2 but you can't necessarily say it's "twice as good" as 2. Also, a poisson model assumes both an infinite support and a that the mean-variance is the same – both are not true with this outcome.

__Response:__ We understand the problem. In the revised manuscript, the severity of CT abnormalities was analyzed by logistic ordinal regression as suggested (function _polr_, R package _MASS_) and the results presented in __Figure E1__. Since, the ordinal models had severe convergence problems while working with the full scale CTSS (0 - 25) and the assumption of odds proportionality (Brant test [@Brant1990]) was severely violated, we decided to define the following severity classes: 0, 1 - 5, 6 - 10 and 11 - 25 CTSS. A shown in __Figure R\@ref(fig:fig-pois-ord-comp)A__ for multi-variable modeling of the CT abnormality severity both Poison and ordinal regression returned comparable results, i.e. in both cases age, sex and acute COVID-19 severity were identified as significant predictors. Still, the ordinal model showed only a limited performance at predicting moderate and severe CT abnormalities (see confusion matrix in __Figure R\@ref(fig:fig-pois-ord-comp)B__), an issue which may be addressed by a broader set of explanatory variables.

## Issue 6

__Issue:__ The word ‘kinetics’ is confusing to me throughout this manuscript. I think you just mean the change or trend of CTTS over time. Given this, I think a repeated measure model is appropriate here. I don’t understand the use of Kendall’s tau the assess this relationship.

__Response:__ We have changed the wording throughout the manuscript to 'change over time' or 'rate of improvement' as suggested by the reviewer and editor. The intention of presenting $\tau$ statistic was to show the reduced drop of CTSS between the consecutive follow-ups, which may be interpreted as decelerated structural lung recovery or chronicity of the residual abnormalities. For the sake of manuscript clarity, we removed the effect size plots from __Figure 5__. Instead, we decided to use more common effect size measure, the non-parametric $r$ statistic for paired data [@Fritz2012] and cite its value and effect size interpretation in the main text section describing the changes of CTSS over time.

## Issue 7

__Issue:__ The figures are much improved, I think there is just too much presented in each figure. All of the extraneous data in the figure can be removed (i.e. the test statistic, df, etc.).

__Response:__ The numeric and text content presented in the Figures was reduced as suggested by the Editor in the manuscript text.

\newpage

# Other changes introduced to the analysis pipeline

## Change 1

The construction of multi-parameter models by backward elimination was accomplished now with another R tool (function _stepAIC()_, package _MASS_ instead of _train()_ function from _caret_). With any of the tools, the same set of significant explanatory variables was included in the final multi-parameter models. The OR and 95$\%$CI estimates are larger due to change of the baseline level for the sex variable as requested by the editor. See: __Figure 3B__.

\newpage

# References

### 2022-02-08 paper/markdown/figure_chunks.Rmd
Paper revision
Modeling strategy changes, figure appearance modified

```{r fig-consort, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Study Flow chart.'}

paper_figures$consort$plot

```

__Figure \@ref(fig:fig-consort). Study Flow chart.__

```{r fig-ct-images, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Non-contrast axial and sagittal chest CTs corresponding to the CT severity score.'}

paper_figures$ct_images$plot

```

__Figure \@ref(fig:fig-ct-images). Non-contrast axial and sagittal chest CTs corresponding to the CT severity score.__

```{r fig-abnormality-risk, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Risk of developing persistent CT abnormalities at the one-year post-COVID-19 follow-up visit.'}

paper_figures$abnormality_risk$plot

```

__Figure \@ref(fig:fig-abnormality-risk). Risk of developing persistent CT abnormalities at the one-year post-COVID-19 follow-up visit.__

```{r fig-ct-images2, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Serial non-contrast axial chest CTs of three study participants with prior COVID-19 pneumonia.'}

paper_figures$ct_images2$plot

```

__Figure \@ref(fig:fig-ct-images2). Serial non-contrast axial chest CTs of three study participants with prior COVID-19 pneumonia.__

```{r fig-ctss-change, fig.width = 7.086614166, fig.height = 8.267716527, fig.cap = 'Change in CT severity score (CTSS) over time.'}

paper_figures$ctss_change$plot

```

__Figure \@ref(fig:fig-ctss-change). Change in CT severity score (CTSS) over time.__

```{r fig-ctss-severity, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Predictors of CT abnormality severity at the 1-year follow-up.'}

suppl_figures$ctss_severity$plot

```

__Figure \@ref(fig:fig-ctss-severity). Predictors of CT abnormality severity at the 1-year follow-up.__

```{r fig-ctss-correlation, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Correlation of CTSS and lung opacity.'}

suppl_figures$ctss_correlation$plot

```

__Figure \@ref(fig:fig-ctss-correlation). Correlation of CTSS and lung opacity.__

```{r fig-age-spline, fig.width = 7.086614166, fig.height = 8.267716527, fig.cap = 'Effects of the age variable splining.'}

rev_figures$age_spline$plot

```

__Figure \@ref(fig:fig-age-spline). Effects of the age variable splining.__

```{r fig-pois-ord-comp, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Comparison of the Poisson and logistic ordinal regression results.'}

rev_figures$pois_ord_comp$plot

```

__Figure \@ref(fig:fig-pois-ord-comp). Comparison of the Poisson and logistic ordinal regression results.__

### 2022-02-08 paper/markdown/figures.Rmd
Paper revision
Modeling strategy changes, figure appearance modified

@@
   flextable() %>%
+  fontsize(size = 10, part = 'all') %>% 
   width(1, width = 3, unit = 'cm') %>% 
@@
   flextable() %>%
-  width(1, width = 3, unit = 'cm') %>% 
-  width(2:5, width = 3.2, unit = 'cm') %>% 
-  width(6, width = 2.5, unit = 'cm') %>% 
+  fontsize(size = 10, part = 'all') %>% 
+  width(1, width = 2.8, unit = 'cm') %>% 
+  width(2:5, width = 2.7, unit = 'cm') %>% 
+  width(6, width = 2.3, unit = 'cm') %>% 
+  width(7, width = 2.3, unit = 'cm') %>% 
   footnote(i = 1, 
@@
 
-```{r fig-consort, fig.width = mm_inch(paper_figures$ct_kinetic$w), fig.height = mm_inch(paper_figures$ct_kinetic$h), dpi = 600, fig.cap = 'Study inclusion flow diagram (placeholder).'}
+```{r fig-consort, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Study Flow chart.'}
 
-ggdraw()
+paper_figures$consort$plot
 
@@
 
-__Figure \@ref(fig:fig-consort). Study inclusion flow diagram (placeholder).__
+__Figure \@ref(fig:fig-consort). Study Flow chart (placeholder).__
 
@@
 
-```{r fig-photos, fig.width = mm_inch(paper_figures$ct_kinetic$w), fig.height = mm_inch(paper_figures$ct_kinetic$h), dpi = 600, fig.cap = 'Representative CT images (placeholder).'}
+```{r fig-ct-images, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Non-contrast axial and sagittal chest CTs corresponding to the CT severity score.'}
 
-ggdraw()
+paper_figures$ct_images$plot
 
@@
 
-__Figure \@ref(fig:fig-photos). Representative CT images (placeholder).__
+__Figure \@ref(fig:fig-ct-images). Non-contrast axial and sagittal chest CTs corresponding to the CT severity score (placeholder).__
 
@@
 
-```{r fig-ct-kinetic, fig.width = mm_inch(paper_figures$ct_kinetic$w), fig.height = mm_inch(paper_figures$ct_kinetic$h), dpi = 600, fig.cap = 'Resolution kinetic of any lung CT abnormalities.'}
+```{r fig-abnormality-risk, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Risk of developing persistent CT abnormalities at the one-year post-COVID-19 follow-up visit.'}
 
-paper_figures$ct_kinetic$plot
+paper_figures$abnormality_risk$plot
 
@@
 
-__Figure \@ref(fig:fig-ct-kinetic). Resolution kinetic of lung CT abnormalities.__
+__Figure \@ref(fig:fig-abnormality-risk). Risk of developing persistent CT abnormalities at the one-year post-COVID-19 follow-up visit.__
 
-Risk of any lung CT abnormalities at the consecutive time points was modeled by mixed-effect logistic regression (random effect: individual, fixed effect: time post COVID-19) in the entire cohort and the acute COVID-19 severity subsets.
-Significance of the fixed time term was determined by likelihood ratio test (LRT) versus the respective random effect-only null model.
+__(A)__ Univariable analysis.
+Risk factors of developing any lung CT abnormalities at the 1-year follow-up visit were identified by logistic regression.
+Odds ratio (OR) significance was determined by Wald Z test.
+OR and with 95$\%$ confidence intervals (CI) are presented in a Forest plot.
+Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.
 
-__(A)__ Percents of individuals with CT abnormalities in the analyzed collectives.
-$\chi$ statistic, degrees of freedom and p values of the LRT test are presented in the plot captions.
-Numbers of individuals with the complete set of consecutive CT scans are shown under the plots.
-
-__(B)__ Recovery between the consecutive follow-ups assessed by Kendall $\tau$ statistic.
+__(B)__ Multivariable analysis.
+Independent risk factors of lung CT abnormalities were identified by multi-parameter logistic regression with backward elimination.
+OR significance was determined by Wald Z test.
+Model prediction was verified by 20-fold cross-validation (CV) and receiver-operating characteristic (ROC).
+OR with 95$\%$ CI are presented in a Forest plot.
+Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.
+Orange: positive correlation, blue: negative correlation, gray: not significant or reference.
+BMI: body mass index, ref.: reference.
 
@@
 
-```{r fig-ctss-kinetic, fig.width = mm_inch(paper_figures$ctss_kinetic$w), fig.height = mm_inch(paper_figures$ctss_kinetic$h), dpi = 600, fig.cap = 'Kinetic of the CT severity score.'}
+```{r fig-ct-images2, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Serial non-contrast axial chest CTs of three study participants with prior COVID-19 pneumonia.'}
 
-paper_figures$ctss_kinetic$plot
+paper_figures$ct_images2$plot
 
@@
 
-__Figure \@ref(fig:fig-ctss-kinetic). Kinetic of the CT severity score.__
-
-CTSS kinetic at the consecutive time points was investigated by Friedman test (grouping by the individual) in the entire cohort and the acute COVID-19 severity subsets.
-The effect size was determined by Kendall W test, differences between particular time points were compared by Wilcoxon test.
-
-__(A)__ Plots display individual CTSS value trajectories as thin gray lines, thick colored lines represent medians, interquartile ranges are presented as colored regions.
-$\chi$ statistic, degrees of freedom and p values of the Friedman test and the Kendall W statistic with 95$\%$ confidence intervals are presented in the plot captions.
-Numbers of individuals with the complete set of consecutive CT scans are shown under the plots.
-
-__(B)__ Recovery between the consecutive follow-ups assessed by Kendall $\tau$ statistic.
+__Figure \@ref(fig:fig-ct-images2). Serial non-contrast axial chest CTs of three study participants with prior COVID-19 pneumonia (placeholder).__
 
@@
 
-```{r fig-pili-risk, fig.width = mm_inch(paper_figures$pili_modeling$w), fig.height = mm_inch(paper_figures$pili_modeling$h), dpi = 600, fig.cap = 'Risk of developing persistent CT abnormalities at the one-year post-COVID-19 follow-up.'}
+```{r fig-ctss-change, fig.width = 7.086614166, fig.height = 8.267716527, fig.cap = 'Change in CT severity score (CTSS) over time.'}
 
-paper_figures$pili_modeling$plot
+paper_figures$ctss_change$plot
 
@@
 
-__Figure \@ref(fig:fig-pili-risk). Risk of developing persistent CT abnormalities at the one-year post-COVID-19 follow-up.__
+__Figure \@ref(fig:fig-ctss-change). Change in CT severity score (CTSS) over time.__
 
-__(A)__ Risk factors of developing any lung CT abnormality at the one-year follow-up visit were identified by a series of logistic models.
-Odds ratio (OR) significance was determined by Wald Z test.
-OR with 95$\%$ confidence intervals are presented in a Forest plot.
-Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.
-
-__(B)__ Independent risk factors of lung CT abnormality were identified by multi-parameter logistic modeling with backward elimination.
-OR significance was determined by Wald Z test.
-Model prediction was verified by 20-fold cross-validation (CV) and receiver-operating characteristic (ROC).
-OR with 95$\%$ confidence intervals are presented in a Forest plot.
-Area under the curve (AUC), sensitivity (Se) and specificity (Sp) for detection of lung CT abnormality in CV are indicated in the plot caption.
-Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.
-
-Orange: positive correlation, blue: negative correlation, gray: not significant, BMI: body mass index, PKY: packages per year, ref.: reference.
-
-\newpage
-
-```{r fig-ctss-model, fig.width = mm_inch(paper_figures$ctss_modeling$w), fig.height = mm_inch(paper_figures$ctss_modeling$h), dpi = 600, fig.cap = 'Predictors of CTSS at the one-year post-COVID-19 follow-up.'}
-
-paper_figures$ctss_modeling$plot
-
-```
-
-__Figure \@ref(fig:fig-ctss-model). Predictors of CTSS at the one-year post-COVID-19 follow-up.__
-
-__(A)__ Factors associated with CTSS at the one-year follow-up visit were identified by a series of Poisson models.
-Regression coefficient ($\beta$) significance was determined by Wald Z test.
-Exponent $\beta$ with 95$\%$ confidence intervals are presented in a Forest plot.
-Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.
-
-__(B)__ Independent factors associated with CTSS were identified by multi-parameter Poisson modeling with backward elimination.
-$\beta$ significance was determined by Wald Z test.
-Model prediction was verified by 20-fold cross-validation (CV).
-Exponent $\beta$ with 95$\%$ confidence intervals are presented in a Forest plot.
-Mean absolute error (MAE) for prediction of CTSS and model R^2^ (Rsq) in CV are indicated in the plot caption.
-Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.
-
-Orange: positive correlation, blue: negative correlation, gray: not significant, BMI: body mass index, PKY: packages per year, ref.: reference.
+CTSS kinetic at the consecutive time points was investigated by Friedman test (grouping by the individual) in the entire cohort and the acute COVID-19 severity subsets.
+The effect size was determined by Kendall W test, differences between particular time points were compared by paired Wilcoxon test.
+Plots display individual CTSS value trajectories as thin gray lines, thick colored lines represent medians, interquartile ranges are presented as colored regions.
+p values of the Friedman test and the Kendall W statistic are presented in the plot captions.
+Numbers of individuals with the complete set of consecutive CT scans are shown under the plots.

### 2022-02-08 paper/markdown/supplementary_material.Rmd
Paper revision
Modeling strategy changes, figure appearance modified

@@
   \usepackage{float} \floatplacement{figure}{H} \floatplacement{table}{H}
-  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}       \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
+  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}       \setcounter{figure}{0} \renewcommand{\thefigure}{E\arabic{figure}}}
 ---
@@
 \beginsupplement
-\newpage
-
-# Supplementary Tables
-
-```{r tab-study-variables, tab.cap = 'Variables available for the analysis. The first 10 table records are shown. The full table is available as a supplementary Excel file.'}
-
-suppl_tables$study_vars[1:10, ] %>% 
-  flextable() %>% 
-  width(1, width = 4, unit = 'cm') %>% 
-  width(2:3, width = 2.5, unit = 'cm') %>% 
-  width(4:5, width = 4, unit = 'cm')
-
-```
-
-\newpage
-
-```{r tab-baseline-severity, tab.cap = 'Baseline characteristic of the acute COVID-19 severity subsets.'}
-
-suppl_tables$baseline_severity %>% 
-  flextable() %>%
-  width(1, width = 1.8, unit = 'cm') %>% 
-  width(2:5, width = 3.9, unit = 'cm') %>% 
-  width(6, width = 2.5, unit = 'cm') %>% 
-  footnote(i = 2:6, 
-           j = 1, 
-           value = as_paragraph('percentage of the acute COVID-19 severity subset (n individuals)'), 
-           part = 'body', 
-           ref_symbols = '1') %>% 
-  footnote(i = 4, 
-           j = 1, 
-           value = as_paragraph('body mass index, kg/m\u00B2, overweight: > 25 kg/m\u00B2, obesity: > 30 kg/m\u00B2'), 
-           part = 'body', 
-           ref_symbols = '2') %>% 
-  footnote(i = 5, 
-           j = 1, 
-           value = as_paragraph('packages per year'), 
-           part = 'body', 
-           ref_symbols = '3') %>% 
-  footnote(i = 1, 
-           j = 6, 
-           value = as_paragraph('Kruskal-Wallis test for differences between the acute COVID-19 severity subsets, ns: not significant'), 
-           part = 'body', 
-           ref_symbols = '4') %>% 
-  footnote(i = 2:6, 
-           j = 6, 
-           value = as_paragraph('\u03C7\u00B2 test for differences between the acute COVID-19 severity subsets, ns: not significant'), 
-           part = 'body', 
-           ref_symbols = '5') %>% 
-  theme_vanilla
-
-```
-
-\newpage
-
-```{r tab-ct-visits, tab.cap = 'Lung CT features in the acute COVID-19 severity subsets at the consecutive follow-up visits.'}
-
-suppl_tables$ct_visits %>% 
-  flextable() %>%
-  width(1, width = 1.8, unit = 'cm') %>% 
-  width(2, width = 2.7, unit = 'cm') %>% 
-  width(3:6, width = 3.2, unit = 'cm') %>% 
-  width(7, width = 2.5, unit = 'cm') %>% 
-  merge_v(1) %>% 
-  footnote(i = 1, 
-           j = 3:6, 
-           value = as_paragraph('percentage of the acute COVID-19 severity subset (n individuals), FUP: follow-up'), 
-           part = 'header', 
-           ref_symbols = '1') %>% 
-  footnote(i = 1, 
-           j = 7, 
-           value = as_paragraph('\u03C7\u00B2 test for differences between the follow-ups, ns: not significant'), 
-           part = 'header', 
-           ref_symbols = '2') %>% 
-  footnote(i = c(6, 21, 36, 51), 
-           j = 2, 
-           value = as_paragraph('ground glass opacity'), 
-           part = 'body', 
-           ref_symbols = '3') %>% 
-  footnote(i = c(11, 26, 41, 56), 
-           j = 2, 
-           value = as_paragraph('acute respiratory distress syndrome'), 
-           part = 'body', 
-           ref_symbols = '4') %>% 
-  footnote(i = c(12, 27, 42, 57), 
-           j = 2, 
-           value = as_paragraph('organizing pneumonia'), 
-           part = 'body', 
-           ref_symbols = '5') %>% 
-  theme_vanilla
-
-```
-
-\newpage
-
-```{r tab-ctss, tab.cap = 'CT severity score at the consecutive visits in the entire cohort and the acute COVID-19 severity subsets.'}
-
-suppl_tables$ctss %>% 
-  flextable() %>%
-  width(1, width = 1.8, unit = 'cm') %>% 
-  width(2:5, width = 3.5, unit = 'cm') %>% 
-  width(6, width = 2.5, unit = 'cm') %>% 
-  footnote(i = 1, 
-           j = 2:5, 
-           value = as_paragraph('FUP: follow-up'), 
-           part = 'header', 
-           ref_symbols = '1') %>% 
-  footnote(i = 1, 
-           j = 6, 
-           value = as_paragraph('Kruskal-Wallis test for differences between the follow-ups, ns: not significant'), 
-           part = 'header', 
-           ref_symbols = '2') %>% 
-  theme_vanilla
-
-```
-
-\newpage
-
-```{r tab-opacity, tab.cap = 'Percentages of opacity and high opacity at the consecutive visits in the entire cohort and the acute COVID-19 severity subsets.'}
-
-suppl_tables$opacity %>% 
-  flextable() %>%
-  width(1, width = 1.8, unit = 'cm') %>% 
-  width(2, width = 1.8, unit = 'cm') %>% 
-  width(3:6, width = 3.5, unit = 'cm') %>% 
-  width(7, width = 2.5, unit = 'cm') %>% 
-  merge_v(1) %>% 
-  footnote(i = 1, 
-           j = 3:6, 
-           value = as_paragraph('FUP: follow-up'), 
-           part = 'header', 
-           ref_symbols = '1') %>% 
-  footnote(i = 1, 
-           j = 7, 
-           value = as_paragraph('Kruskal-Wallis test for differences between the follow-ups, ns: not significant'), 
-           part = 'header', 
-           ref_symbols = '2') %>% 
-  theme_vanilla
-
-```
 
@@
 
-```{r fig-ct-sev, fig.width = mm_inch(suppl_figures$ct_kinetic$w), fig.height = mm_inch(suppl_figures$ct_kinetic$h), dpi = 600, fig.cap = 'Resolution kinetic of moderate-to-severe lung CT abnormalities.'}
+```{r fig-ctss-severity, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Predictors of CT abnormality severity at the 1-year follow-up.'}
 
-suppl_figures$ct_kinetic$plot
+suppl_figures$ctss_severity$plot
 
@@
 
-__Figure \@ref(fig:fig-ct-sev). Resolution kinetic of moderate-to-severe lung CT abnormalities.__
+__Figure E\@ref(fig:fig-ctss-severity). Predictors of CT abnormality severity at the 1-year follow-up.__
 
-Risk of moderate-to-severe lung CT abnormalities (CT severity score > 5) at the consecutive time points was modeled by mixed-effect logistic regression (random effect: individual, fixed effect: time post COVID-19) in the entire cohort and the acute COVID-19 severity subsets.
-Significance of the fixed time term was determined by likelihood ratio test (LRT) versus the respective random effect-only null model.
+__(A)__ Univariable analysis.
+Factors associated with CT abnormality severity (classes: 0, 1 - 5, 6 - 10, 11 - 25 CT severity score [CTSS] points) were identified by ordinal logistic regression.
+Odds ratio (OR) significance was determined by Wald Z test.
+OR and with 95$\%$ confidence intervals (CI) are presented in a Forest plot.
+Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.
 
-__(A)__ Percents of individuals with moderate-to-severe CT abnormalities in the analyzed collectives.
-$\chi$ statistic, degrees of freedom and p values of the LRT test are presented in the plot captions.
-Numbers of individuals with the complete set of consecutive CT scans are shown under the plots.
-
-__(B)__ Recovery between the consecutive follow-ups assessed by Kendall $\tau$ statistic.
+__(B)__ Multivariable analysis.
+Independent factors associated with CT abnormality severity (classes: 0, 1 - 5, 6 - 10, 11 - 25 CTSS) were identified by ordinal logistic regression with backward elimination.
+OR significance was determined by Wald Z test.
+Model prediction was verified by 20-fold cross-validation (CV), miss-classification error and Cohen's $\kappa$ statistic.
+OR with 95$\%$ CI are presented in a Forest plot.
+Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.
+Orange: positive correlation, blue: negative correlation, gray: not significant or reference.
+BMI: body mass index, ref.: reference.
 
@@
 
-```{r fig-ct-opacity, fig.width = mm_inch(suppl_figures$ct_opacity$w), fig.height = mm_inch(suppl_figures$ct_opacity$h), dpi = 600, fig.cap = 'Resolution kinetic of lung CT opacity.'}
+```{r fig-ctss-correlation, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Correlation of CTSS and lung opacity.'}
 
-suppl_figures$ct_opacity$plot
+suppl_figures$ctss_correlation$plot
 
@@
 
-__Figure \@ref(fig:fig-ct-opacity). Resolution kinetic of lung CT opacity.__
-
-Risk of any lung CT opacity at the consecutive time points was modeled by mixed-effect logistic regression (random effect: individual, fixed effect: time post COVID-19) in the entire cohort and the acute COVID-19 severity subsets.
-Significance of the fixed time term was determined by likelihood ratio test (LRT) versus the respective random effect-only null model.
-
-__(A)__ Percents of individuals with CT opacity in the analyzed collectives.
-$\chi$ statistic, degrees of freedom and p values of the LRT test are presented in the plot captions.
-Numbers of individuals with the complete set of consecutive CT scans are shown under the plots.
-
-__(B)__ Recovery between the consecutive follow-ups assessed by Kendall $\tau$ statistic.
-
-\newpage
-
-```{r fig-opacity-kinetic, fig.width = mm_inch(suppl_figures$opacity_kinetic$w), fig.height = mm_inch(suppl_figures$opacity_kinetic$h), dpi = 600, fig.cap = 'Kinetic of the lung opacity.'}
-
-suppl_figures$opacity_kinetic$plot
-
-```
-
-__Figure \@ref(fig:fig-opacity-kinetic). Kinetic of the lung opacity.__
-
-Kinetic of the lung opacity at the consecutive time points was investigated by Friedman test (grouping by the individual) in the entire cohort and the acute COVID-19 severity subsets.
-The effect size was determined by Kendall W test, differences between particular time points were compared by Wilcoxon test.
-
-__(A)__ Plots display individual opacity trajectories as thin gray lines, thick colored lines represent medians, interquartile ranges are presented as colored regions.
-$\chi$ statistic, degrees of freedom and p values of the Friedman test and the Kendall W statistic with 95$\%$ confidence intervals are presented in the plot captions.
-Numbers of individuals with the complete set of consecutive CT scans are shown under the plots.
-
-__(B)__ Recovery between the consecutive follow-ups assessed by Kendall $\tau$ statistic.
-
-\newpage
-
-```{r fig-opacity-risk, fig.width = mm_inch(suppl_figures$opacity_risk$w), fig.height = mm_inch(suppl_figures$opacity_risk$h), dpi = 600, fig.cap = 'Risk of developing lung opacity at the one-year post-COVID-19 follow-up.'}
-
-suppl_figures$opacity_risk$plot
-
-```
-
-__Figure \@ref(fig:fig-opacity-risk). Risk of developing lung opacity at the one-year post-COVID-19 follow-up.__
-
-__(A)__ Risk factors of developing opacity at the one-year follow-up visit were identified by a series of logistic models.
-Odds ratio (OR) significance was determined by Wald Z test.
-OR with 95$\%$ confidence intervals are presented in a Forest plot.
-Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.
-
-__(B)__ Independent risk factors of opacity were identified by multi-parameter logistic modeling with backward elimination.
-OR significance was determined by Wald Z test.
-Model prediction was verified by 20-fold cross-validation (CV) and receiver-operating characteristic (ROC).
-OR with 95$\%$ confidence intervals are presented in a Forest plot.
-Area under the curve (AUC), sensitivity (Se) and specificity (Sp) for detection of opacity in CV are indicated in the plot caption.
-Numbers of complete observations and the reference levels of the explanatory variables are indicated in the Y axis.
-
-Orange: positive correlation, blue: negative correlation, gray: not significant, BMI: body mass index, PKY: packages per year, ref.: reference.
-
-\newpage
-
-```{r fig-opacity-rater, fig.width = mm_inch(suppl_figures$any_abnormality_rater$w), fig.height = mm_inch(suppl_figures$any_abnormality_rater$h), dpi = 600, fig.cap = 'Performance of automated opacity assessment at detection of lung CT abnormalities.'}
-
-suppl_figures$any_abnormality_rater$plot
-
-```
-
-__Figure \@ref(fig:fig-opacity-rater). Performance of automated opacity assessment at detection of lung CT abnormalities.__
-
-Lung CT abnormality were defined either by the radiologist or by the software as opacity $\geq$ 1$\%$ lung volume.
-Performance of the opacity at detecting CT abnormalities was investigated at the consecutive follow-ups by receiver-operating characteristic (ROC) and Cohen's $\kappa$ statistic.
-Numbers of complete observations and CT abnormalities (events) are indicated under the plots.
-
-__(A)__ ROC curves annotated with sensitivity (Se), specificity (Sp) and area under the curve (AUC) values with 95$\%$ confidence intervals.
-
-__(B)__ $\kappa$ with 95$\%$ confidence intervals presented in a forest plot.
-
-\newpage
-
-```{r fig-ggo-rater, fig.width = mm_inch(suppl_figures$ggo_rater$w), fig.height = mm_inch(suppl_figures$ggo_rater$h), dpi = 600, fig.cap = 'Performance of automated opacity assessment at detection of GGO.'}
-
-suppl_figures$ggo_rater$plot
-
-```
-
-__Figure \@ref(fig:fig-ggo-rater). Performance of automated opacity assessment at detection of GGO.__
-
-Lung GGO were defined either by the radiologist or by the software as opacity $\geq$ 1$\%$ lung volume.
-Performance of the opacity at detecting GGO was investigated at the consecutive follow-ups by receiver-operating characteristic (ROC) and Cohen's $\kappa$ statistic.
-Numbers of complete observations and CT abnormalities (events) are indicated under the plots.
-
-__(A)__ ROC curves annotated with sensitivity (Se), specificity (Sp) and area under the curve (AUC) values with 95$\%$ confidence intervals.
-
-__(B)__ $\kappa$ with 95$\%$ confidence intervals presented in a forest plot.
-
-\newpage
-
-```{r fig-ctss-opacity, fig.width = mm_inch(suppl_figures$ctss_opacity$w), fig.height = mm_inch(suppl_figures$ctss_opacity$h), dpi = 600, fig.cap = 'Correlation of CTSS and lung opacity.'}
-
-suppl_figures$ctss_opacity$plot
-
-```
-
-__Figure \@ref(fig:fig-ctss-opacity). Correlation of CTSS and lung opacity.__
+__Figure E\@ref(fig:fig-ctss-correlation). Correlation of CTSS and lung opacity.__
 
@@
 Points represent single observations, blue lines with gray regions represent fitted linear trend with 95$\%$ confidence interval.
-Correlation coefficients (r), p values and numbers of complete observations are indicated in the plot captions.
+Correlation coefficients ($\rho$), p values and numbers of complete observations are indicated in the plot captions.

### 2022-02-08 analysis scripts/distribution.R

@@ -10,130 +10,80 @@
   
   insert_msg('Globals setup')
   
-  distr$responses <- c('ctss', 'perc_opac')
+  distr$responses <- c('ctss', 'perc_opac', 'perc_hiopac', 'log_ctss', 'log_perc_opac', 'log_perc_hiopac')
   
-# Normality of the CT severity score, percent opacity and percent high opacity at the time points -----
+# Normality of the CT severity score, percent opacity and percent high opacity at the time points: log and identity -----
   
   insert_msg('CT severity score normality check')
   
   ## normality in the visit groups
   
-  distr$ctss_normality$time_test_ident <- radio$long %>% 
-    dlply(.(visit)) %>% 
-    map(~shapiro_test(.x, ctss, perc_opac)) %>% 
-    map2_dfr(., names(.), ~mutate(.x, visit = .y))
-
-  distr$ctss_normality$time_plot_ident$ctss <- radio$long %>% 
-    dlply(.(visit)) %>% 
-    map(~ggplot(.x, aes(sample = ctss)) + 
-          geom_qq() + 
-          geom_qq_line())
-  
-  distr$ctss_normality$time_plot_ident$perc_opac <- radio$long %>% 
-    dlply(.(visit)) %>% 
-    map(~ggplot(.x, aes(sample = perc_opac)) + 
-          geom_qq() + 
-          geom_qq_line())
-  
-  ## normality in the time and time:severity groups
-  
-  distr$ctss_normality$time_sev_test_ident <- radio$long %>% 
-    dlply(.(visit, severity)) %>% 
-    map(~shapiro_test(.x, ctss, perc_opac)) %>% 
+  distr$ctss_normality$time_test_visit <- radio$long %>% 
+    mutate(log_ctss = log(ctss + 1), 
+           log_perc_opac = log(perc_opac + 1), 
+           log_perc_hiopac = log(perc_hiopac + 1)) %>% 
+    explore(split_factor = 'visit', 
+            variables = distr$responses, 
+            what = 'normality') %>% 
     map2_dfr(., names(.), ~mutate(.x, visit = .y))
   
-  distr$ctss_normality$time_sev_plot_ident$ctss <- radio$long %>% 
-    dlply(.(visit, severity)) %>% 
-    map(~ggplot(.x, aes(sample = ctss)) + 
-          geom_qq() + 
-          geom_qq_line())
+  distr$ctss_normality$time_plot_visit <- radio$long %>% 
+    mutate(log_ctss = log(ctss + 1), 
[…]

### 2022-02-08 analysis scripts/eda.R

@@ -3,137 +3,135 @@
 # statistics of the CT severity score are calculated
 
   insert_head()
+
+# container list and globals ------
+
+  ed_analysis <- list()
   
-# container list ------
+  ed_analysis$variables <-  c('age', 'age_60', 'bmi_class', 'pky', 'pky_class', 'smoking')
   
-  eda <- list()
+  ed_analysis$test_types <- c('anova', 'chisq_test', 'chisq_test', 'kruskal_test', 'chisq_test', 'chisq_test')
   
+# distribution of the age and pky variable in the severity groups ------
+  
+  insert_msg('Distributon and variance homogeneity for age')
+  
+  ed_analysis$age$normality <- explore(data = radio$clear, 
+                                       variables = c('age', 'pky'), 
+                                       split_factor = 'severity', 
+                                       what = 'normality') %>% 
+    map2_dfr(., names(.), ~mutate(.x, severity = .y))
+  
+  ed_analysis$age$homogeneity <- compare_variables(radio$clear, 
+                                                   variables = c('age', 'pky'), 
+                                                   split_factor = 'severity', 
+                                                   what = 'variance')
+
 # baseline clinical characteristic -----
-  
+
   insert_msg('Baseline clinics, entire cohort')
-  
-  ## analyses
 
-  eda$baseline$analyses_cohort <- c('age', radio$long_expl_var) %>% 
-    map(analyze_feature, 
-        inp_tbl = rbind(mutate(radio$clear, dummy_var = factor('dummy1')), 
-                        mutate(radio$clear, dummy_var = factor('dummy2'))), 
-        split_var = 'dummy_var') %>% 
-    set_names(c('age', radio$long_expl_var))
-
-  ## summary table
-  
-  eda$baseline$summary_cohort <- eda$baseline$analyses_cohort %>% 
-    map_dfr(get_feature_summary) %>% 
-    mutate(label = translate_var(variable), 
-           statistic = dummy1) %>% 
-    select(variable, label, statistic)
+  ed_analysis$baseline <- explore(data = radio$clear, 
+                                  variables = c(ed_analysis$variables, 'severity'), 
+                                  what = 'table', 
+                                  pub_styled = TRUE, 
+                                  simplify_p = TRUE)
   
 # baseline characteristic in the severity groups -----
   
   insert_msg('Baseline clinics, severity groups')
   
-  ## analyses
+  ## descriptive statistics
   
-  eda$baseline$analyses_severity <- c('age', radio$long_expl_var) %>% 
[…]

### 2022-02-08 analysis scripts/interrater.R

@@ -27,31 +27,38 @@
   
   insert_msg('Correlations CTSS versus opacity')
   
+  ## analysis table
+  
+  rater$ctss$analysis_tbl <- radio$long %>% 
+    rater_tbl(ctss, perc_opac)
+  
   ## analyses
   
-  rater$ctss$analyses <- radio$long %>% 
-    rater_tbl(ctss, perc_opac) %>% 
-    map(cor_test, 
-        ctss, 
-        perc_opac, 
-        method = 'spearman') %>% 
-    map2_dfr(., names(.), ~mutate(.x, visit = .y)) %>% 
-    mutate(n_complete = radio$long %>% 
-             rater_tbl(ctss, perc_opac) %>% 
-             map(nrow), 
-           plot_lab = paste0('r = ', signif(cor, 2), ', ', format_p(p), ', n = ', n_complete))
+  rater$ctss$analyses <- correlate_variables(!!!rater$ctss$analysis_tbl, 
+                                             variables = c('ctss', 'perc_opac'), 
+                                             type = 'spearman', 
+                                             boot_method = 'bca', 
+                                             pub_styled = TRUE, 
+                                             simplify_p = TRUE) %>% 
+    mutate(eff_size = stri_replace(eff_size, fixed = 'rho', replacement = '\u03C1'),
+           eff_size = stri_replace_all(eff_size, fixed = '0.', replacement = '.'), 
+           significance = stri_replace(significance, fixed = '0.', replacement = '.'), 
+           plot_caption = paste0(eff_size, ', ', significance, ', n = ', n, sep = ', '), 
+           plot_caption = stri_replace(plot_caption, regex = '\\[.*\\]', replacement = ''))
+
 
   ## plots
   
-  rater$ctss$plots <- list(data = radio$long %>% 
-                             rater_tbl(ctss, perc_opac), 
+  rater$ctss$plots <- list(data = rater$ctss$analysis_tbl, 
                            plot_title = globals$visit_labels, 
-                           fill_color = globals$visit_colors, 
-                           plot_subtitle = rater$ctss$analyses$plot_lab) %>% 
-    pmap(correlation_plot, 
-         jitter_w = 0.1) %>% 
+                           point_color = globals$visit_colors, 
+                           plot_subtitle = rater$ctss$analyses$plot_caption) %>% 
+    pmap(plot_correlation, 
+         variables = c('ctss', 'perc_opac'), 
[…]

### 2022-02-08 analysis scripts/kinetic.R

@@ -1,7 +1,5 @@
-# This script 'models' kinetic of the CT score in the entire cohort and severity groups
+# This script 'models' changes of the CT score and percent opacity in time in the entire cohort and severity groups
 # by Friedman test.
-# Risk of incomplete resolution of any, moderate-to-severe, severe and particular type abnormalities
-# is estimated by mixed-effect logistic modeling. Only the participants with the complete longitudinal CT scan sets are analyzed
 
   insert_head()
   
@@ -11,220 +9,99 @@
   
 # analysis table ----
   
+  insert_msg('Analysis table')
+  
   kinetic$analysis_tbl <- radio$long %>% 
     filter(ID %in% radio$complete_long_ids)
   
   kinetic$analysis_tbl <- kinetic$analysis_tbl %>% 
-    dlply(.(severity)) %>% 
+    dlply(.(severity), as_tibble) %>% 
     c(list(cohort = kinetic$analysis_tbl), .)
   
-# CTSS, opacity and hig opacity percentage kinetics -----
+  kinetic$analysis_tbl <- kinetic$analysis_tbl %>% 
+    map(select, ID, severity, visit, time_months, ctss, perc_opac)
   
-  insert_msg('CTSS kinetic')
+# Friedman test -----
   
-  ## test
-  
-  kinetic$numeric$analysis <- list(ctss = ctss ~ visit|ID, 
-                                   perc_opac = perc_opac ~ visit|ID, 
-                                   perc_hiopac = perc_hiopac ~ visit|ID) %>% 
-    map(function(outcome) kinetic$analysis_tbl %>% 
-          map(friedman_test, 
-                  formula = outcome) %>% 
-          map2_dfr(., names(.), ~mutate(.x, subset = .y)))
-
-  ## summary
-  
-  kinetic$numeric$summary <- kinetic$numeric$analysis %>% 
-    map(mutate, 
-        p_value = p, 
-        p_adj = p.adjust(p_value, 'BH'), 
-        plot_caption = paste0('\u03C7 = ', signif(statistic, 2), 
-                              ', df = 3, ', format_p(p_value, 2)))
-  
-  ## effect size determined by Kendall W, updating the plot_caption
+  insert_msg('Friedman test')
   
   plan('multisession')
   
-  kinetic$numeric$effsize <- list(ctss = ctss ~ visit|ID, 
-                                  perc_opac = perc_opac ~ visit|ID, 
[…]

### 2022-02-08 analysis scripts/risk_modeling.R

@@ -1,0 +1,183 @@
+# This script models risk of any abnormality, any opacity, GGO and reticulation,  
+# (logistic regression) at the 1-year follow-up. Explanatory variables
+# are sex, age, severity, BMI, smoking, PKY. Multi-parameter models are constructed by backward elimination.
+
+  insert_head()
+  
+# Container list -----
+  
+  risk <- list()
+  
+# Globals ------
+  
+  insert_msg('Globals')
+
+  risk$outcomes <-  c('ggo_fup4', 
+                      'retic_fup4', 
+                      'ctss_any_fup4', 
+                      'opacity_fup4')
+  
+  ## baseline levels
+  
+  risk$base_levels <- radio$long_expl_var %>% 
+    map_chr(~levels(radio$clear[[.x]])[1]) %>% 
+    set_names(radio$long_expl_var)
+
+# Serial univariate modeling -----
+
+  insert_msg('Serial univariate modeling')  
+  
+  ## modeling objects
+  
+  risk$univariate$models <- risk$outcomes %>% 
+    map(function(resp) radio$long_expl_var %>% 
+          map(~make_lm(data = radio$clear, 
+                       response = resp, 
+                       indep_variable = .x, 
+                       mod_fun = glm, 
+                       family = 'binomial')) %>% 
+          set_names(radio$long_expl_var)) %>% 
+    set_names(risk$outcomes)
+  
+  ## residuals: normality and variance homogeneity testing
+
+  risk$univariate$resid_normality_eov <- risk$univariate$models %>% 
+    map(~map(.x, summary, type = 'assumptions') %>% 
+          map2_dfr(., names(.), ~mutate(.x, variable = .y))) %>% 
+    map2_dfr(., names(.), ~mutate(.x, response = .y))
+
+  ## diagnostic plots: residuals
+  
+  risk$univariate$diagnostic_plots <- risk$univariate$models %>% 
+    map(~map(.x, plot, cust_theme = globals$common_theme))
+  
+  ## fit statistics
+  
+  risk$univariate$fit_stats <- risk$univariate$models %>% 
+    map(~map(.x, summary, type = 'fit') %>% 
+          map2_dfr(., names(.), ~mutate(.x, variable = .y)))
+  
+  ## inference summary, appending with the baseline levels
+  
+  risk$univariate$inference <- risk$univariate$models %>% 
+    map(~map(.x, summary, type = 'inference', transf_fun = exp) %>% 
+          map(~mutate(.x, 
+                      plot_order = 1:nrow(.x), 
[…]

### 2022-02-08 analysis scripts/risk_modeling_gam.R

@@ -1,0 +1,227 @@
+# Models risk of any CT abnormalities, any opacity, GGO and reticulations with logistic GAM.
+# The numeric variables: age and pack-years are treated with cubic splines.
+
+  insert_head()
+  
+# container list ----
+  
+  risk_gam <- list()
+  
+# Globals ------
+  
+  insert_msg('Globals')
+  
+  risk_gam$outcomes <-  c('ggo_fup4', 
+                          'retic_fup4', 
+                          'ctss_any_fup4', 
+                          'opacity_fup4')
+  
+  ## baseline levels
+  
+  risk_gam$base_levels <- c('sex', 'bmi_class', 'smoking', 'severity') %>% 
+    map_chr(~levels(radio$clear[[.x]])[1]) %>% 
+    set_names(c('sex', 'bmi_class', 'smoking', 'severity'))
+  
+  ## univariate modeling formulas
+  
+  risk_gam$uni_forms <- risk_gam$outcomes %>% 
+    map(function(resp) c('s(age, bs = "tp", k = 20)', 's(pky, bs = "tp", k = 15)') %>% 
+          map(~paste(resp, .x, sep = '~')) %>%
+          map(as.formula) %>%
+          set_names(c('age', 'pky'))) %>% 
+    set_names(risk_gam$outcomes)
+  
+  ## multi-variate modeling formulas
+  
+  risk_gam$multi_forms <- risk_gam$outcomes %>% 
+    map(~paste(.x, '~ sex + severity + s(age, bs = "tp", k = 20)')) %>% 
+    map(as.formula) %>% 
+    set_names(risk_gam$outcomes)
+  
+# Univariate modeling -----
+  
+  insert_msg('Univariate modeling')
+  
+  ## model objects
+  
+  risk_gam$univariate$models <- risk_gam$uni_forms  %>% 
+    map(~map(.x, 
+             ~make_lm(formula = .x, 
+                      data = radio$clear, 
+                      family = 'binomial', 
+                      mod_fun = gam, 
+                      method = 'REML')))
+  
+
+  ## residuals, normality testing
+  
+  risk_gam$univariate$resid_normality_eov <- risk_gam$univariate$models %>% 
+    map(~map(.x, summary, type = 'assumptions') %>% 
+          map2_dfr(., names(.), ~mutate(.x, variable = .y))) %>% 
+    map2_dfr(., names(.), ~mutate(.x, response = .y))
+  
[…]

### 2022-02-08 analysis scripts/score_modeling.R

@@ -1,0 +1,234 @@
+# This script models CTSS (Poisson regression) and CTSS class,  
+# (logistic regression) at the 1-year follow-up. Explanatory variables
+# are sex, age, severity, BMI, smoking, PKY. Multi-parameter models are constructed by backward elimination.
+
+  insert_head()
+  
+# Container list -----
+  
+  score <- list()
+
+# Globals ------
+  
+  insert_msg('Globals')
+
+  score$outcomes <-  c('ctss_fup4', 'ctss_class_fup4')
+
+  ## baseline and baseline n numbers
+  
+  score$base_levels <- radio$long_expl_var %>% 
+    map_chr(~levels(radio$clear[[.x]])[1]) %>% 
+    set_names(radio$long_expl_var)
+
+  ## faceting formulas 
+  
+  score$facet_forms <- paste('.~', radio$long_expl_var) %>% 
+    map(as.formula)
+  
+# Serial univariate modeling -----
+
+  insert_msg('Serial univariate modeling')  
+  
+  ## modeling objects
+  
+  score$univariate$models <- list(response  = score$outcomes, 
+                                  mod_fun = list(glm, polr), 
+                                  family = list('poisson', NULL)) %>% 
+    pmap(function(response, mod_fun, family) radio$long_expl_var %>% 
+           map(~make_lm(data = radio$clear, 
+                        response = response, 
+                        indep_variable = .x, 
+                        mod_fun = mod_fun, 
+                        family = family)) %>% 
+           set_names(radio$long_expl_var)) %>% 
+    set_names(score$outcomes)
+
+  ## residuals: normality and variance homogeneity testing
+  
+  score$univariate$resid_normality_eov <- score$univariate$models %>% 
+    map(~map(.x, summary, type = 'assumptions') %>% 
+          map2_dfr(., names(.), ~mutate(.x, variable = .y))) %>% 
+    map2_dfr(., names(.), ~mutate(.x, response = .y))
+
+  ## diagnostic plots: residuals
+  
+  score$univariate$diagnostic_plots <- score$univariate$models %>% 
+    map(~map(.x, plot, cust_theme = globals$common_theme))
+  
+  ## diagnostic plots: response distribution
+  
+  score$univariate$distr_plots <- map2(score$univariate$models, 
+                                       c(10, 4), 
+                                       function(models, bin_no) list(lm_analysis_object = models, 
+                                                                     facet_formula = score$facet_forms) %>% 
+                                         pmap(plot, 
[…]

### 2022-02-08 data_analysis.R

@@ -4,31 +4,39 @@
   
 # tools -----
   
-  c('./tools/sys_tools.R', 
-    './tools/project_globals.R', 
-    './tools/counting_tools.R', 
-    './tools/modeling_tools.R', 
-    './tools/lm_qc_tools.R', 
-    './tools/roc_toolbox.R', 
-    './tools/project_tools.R') %>% 
-    walk(source)
-  
+  library(soucer) ## available from https://github.com/PiotrTymoszuk/soucer
   library(DescTools)
   library(rstatix)
   library(ggpubr)
   library(furrr)
   library(caret)
+  library(doParallel)
   library(vcd)
+  library(MASS)
+  library(broom)
+  library(mgcv)
+  library(exda) ## available from https://github.com/PiotrTymoszuk/ExDA
+  library(lmqc) ## available from https://github.com/PiotrTymoszuk/lmqc
+  library(caretExtra) ## available from https://github.com/PiotrTymoszuk/caretExtra
+  
+  select <- dplyr::select
+  
+  c('./tools/roc_toolbox.R', 
+    './tools/project_globals.R', 
+    './tools/project_tools.R') %>% 
+  source_all(message = TRUE, crash = TRUE)
   
 # analysis scripts -----
   
   c('./analysis scripts/distribution.R', ## distribution checks
-    './analysis scripts/eda.R', ## explorative data analysis
-    './analysis scripts/kinetic.R', ## kinetic analysis
-    './analysis scripts/risk_score.R', ## risk and score modeling
-    './analysis scripts/interrater.R') %>% ## inter-rater comparison of the CTSS and opacity
-    walk(source)
-  
+     './analysis scripts/eda.R', ## explorative data analysis
+     './analysis scripts/kinetic.R', ## kinetic analysis
+    './analysis scripts/interrater.R', ## inter-rater comparison of the CTSS and opacity
+    './analysis scripts/risk_modeling.R', ## abnormality risk modeling
+    './analysis scripts/score_modeling.R', ## CTSS modeling
+    './analysis scripts/risk_modeling_gam.R') %>% ## GAM modeling and comparison with the age-cutoff models
+    source_all(message = TRUE, crash = TRUE) 
+
 # END -----
   
   insert_tail()

### 2022-02-08 data_import.R

@@ -7,12 +7,12 @@
   library(foreign)
   library(stringi)
   library(readxl)
+  library(soucer) ## available from https://github.com/PiotrTymoszuk/soucer
 
-  c('./tools/sys_tools.R', 
-    './tools/project_globals.R', 
+  c('./tools/project_globals.R', 
     './tools/project_tools.R') %>% 
-    walk(source)
-  
+    source_all(message = TRUE, crash = TRUE)
+      
   insert_head()
   
 # container list ----
@@ -68,7 +68,7 @@
   
   radio$clear <- radio$clear %>% 
     mutate(sex = car::recode(sex, "'m' = 'male'; 'w' = 'female'"), 
-           sex = factor(sex, c('male', 'female')), 
+           sex = factor(sex, c('female', 'male')), 
            age = date_fup4 - date_birth - 1, 
            age = round(as.numeric(age, format = 'days')/365.25), 
            age_60 = cut(age, c(-Inf, 60, Inf), c('up to 60', '>60')), 
@@ -127,7 +127,7 @@
     if(is.character(radio$long[[i]])) {
       
       radio$long <- radio$long %>% 
-        mutate(!!sym(i) := factor(.data[[i]]))
+        mutate(!!i := factor(.data[[i]]))
       
     }
     

### 2022-02-08 exec_all.R

@@ -1,60 +1,17 @@
 # This program executes the project scripts
 
-# libraries -----
-
-  require(readr)
+  library(soucer)
 
 # listing scripts ----
 
-  executable_scripts <- c('data_import.R', 
-                          'data_analysis.R', 
-                          'render_paper.R') 
+  exec_log <- source_all(c('data_import.R', 
+                           'data_analysis.R', 
+                           'render_paper.R'), 
+                         message = TRUE, crash = FALSE)
+  
+  print(exec_log)
 
-# error container ----
-  
-  exec_errors <- list()
-  
-# executing the script list ----
-  
-  for(file in executable_scripts) {
-    
-    diagn_output <- try(source(file, encoding = 'UTF-8'), silent = T) ## UTF-8 encoding cause of greek letters 
-                                                                      ## in plots and tables
-    
-    if(class(diagn_output) == 'try-error') {
-      
-      exec_errors <- c(exec_errors, 'execution failed')
-      
-    } else {
-      
-      exec_errors <- c(exec_errors, 'execution successful')
-      
-    }
-
-  }
-  
-  require(dplyr)
-  
-  exec_errors <- exec_errors %>% 
-    reduce(c) %>% 
-    tibble(execution_result = .)
-  
-  exec_errors <- exec_errors %>% 
-    mutate(file = executable_scripts) %>% 
-    select(file, execution_result)
-  
-  rm(diagn_output, executable_scripts)
-  
-  message(rep('-', 60))
-  print(exec_errors)
-  message(rep('-', 60))
-  
-# saving execution results on the disc ----
-
-  exec_errors %>% 
-    write_tsv('exec_log.log')
-  
-# saving the image -----
+  'render_paper.R'
   
   save.image()
   

### 2022-02-08 paper scripts/deploy_paper.R

@@ -1,39 +1,54 @@
 # This script renders the legends for tables and figures ----
 
-# data and tools ----
-
-  source('./tools/sys_tools.R')
-
-  library(knitr)
-  library(bookdown)
-  library(kableExtra)
-  library(rmarkdown)
-  library(flextable)
-
   insert_head()
+
+# Exporting the figure chunks -----
   
-# Specific functions -----
+  insert_msg('Figure chunk export')
   
-  mm_inch <- function(input_mm) {
-    
-    return(0.0393700787 * input_mm)
-    
-  }
+  insert_figure(paper_figures$consort, 
+                paper_figures$ct_images, 
+                paper_figures$abnormality_risk, 
+                paper_figures$ct_images2, 
+                paper_figures$ctss_change, 
+                suppl_figures$ctss_severity, 
+                suppl_figures$ctss_correlation, 
+                rev_figures$age_spline, 
+                rev_figures$pois_ord_comp, 
+                file = './paper/markdown/figure_chunks.Rmd', 
+                ref_names = stri_replace_all(c(names(paper_figures), 
+                                               names(suppl_figures), 
+                                               names(rev_figures)), 
+                                             fixed = '_', 
+                                             replacement = '-'), 
+                captions = c('Study Flow chart.', 
+                             'Non-contrast axial and sagittal chest CTs corresponding to the CT severity score.', 
+                             'Risk of developing persistent CT abnormalities at the one-year post-COVID-19 follow-up visit.', 
+                             'Serial non-contrast axial chest CTs of three study participants with prior COVID-19 pneumonia.', 
+                             'Change in CT severity score (CTSS) over time.', 
+                             'Predictors of CT abnormality severity at the 1-year follow-up.', 
+                             'Correlation of CTSS and lung opacity.', 
+                             'Effects of the age variable splining.', 
+                             'Comparison of the Poisson and logistic ordinal regression results.'), 
+                append = FALSE, 
+                add_extern_legend = TRUE)
 
 # Generating figure legend file -----
   
   insert_msg('Rendering figure and table legend file')
   
   render('./paper/markdown/figures.Rmd', 
-         output_format = word_document2(number_sections = F, theme = 'readable'), 
+         output_format = word_document2(number_sections = FALSE, theme = 'readable'), 
          output_dir = './paper/')
   
   render('./paper/markdown/supplementary_material.Rmd', 
[…]

### 2022-02-08 paper scripts/paper_figures.R

@@ -6,322 +6,175 @@
 
   paper_figures <- list()
   suppl_figures <- list()
+  rev_figures <- list() ## response to the statistical reviewer
   
 # Figure 1: Consort plot - placeholder, provided by the Author team ----
   
   insert_msg('Figure 1: CONSORT, provided by the authors')
   
+  paper_figures$consort <- ggdraw() %>% 
+    as_figure(label = 'figure_1_consort_placeholder', 
+              w = 180, 
+              h = 180)
+  
 # Figure 2: CT images, provided by the Author team -----
   
   insert_msg('Figure 2: CT images, provided by the authors')
   
-# Figure 3: kinetics of any CT abnormalities -----
+  paper_figures$ct_images <- ggdraw() %>% 
+    as_figure(label = 'figure_2_ct_placeholder', 
+              w = 180, 
+              h = 180)
   
-  insert_msg('Figure 3: kinetic of any CT abnormalities and CTSS')
+# Figure 3: modeling of the CT abnormality risk -----
   
-  paper_figures$ct_kinetic$upper_panel <- kinetic$binary$plots[c('cohort.ctss_any', 
-                                                                 'mild.ctss_any', 
-                                                                 'moderate.ctss_any', 
-                                                                 'severe.ctss_any', 
-                                                                 'critical.ctss_any')] %>% 
-    map2(., globals$subset_labels, 
-         ~.x + 
-           labs(title = .y) + 
-           scale_y_continuous(limits = c(0, 115), 
-                              breaks = seq(0, 100, by = 25)) + 
-           scale_x_continuous(limits = c(2, 13), 
-                              breaks = c(2, 3, 6, 12))) %>% 
-    plot_grid(plotlist = ., 
-              ncol = 3, 
-              align = 'hv', 
-              axis = 'tblr')
-  
-  paper_figures$ct_kinetic$bottom_panel <- kinetic$binary$eff_plots$ctss_any + 
-    theme(legend.position = 'none') + 
-    facet_grid(.~severity, 
-               labeller = as_labeller(globals$subset_labels))
+  insert_msg('Figure 3: CT abnormality risk')
 
+  paper_figures$abnormality_risk <-  plot_grid(risk$univariate$plots$ctss_any_fup4, 
+                                               plot_grid(risk$multivariate$plots$ctss_any_fup4, 
+                                                         ggdraw(), 
+                                                         nrow = 2, 
+                                                         rel_heights = c(0.6, 0.4)), 
[…]

### 2022-02-08 paper scripts/paper_tables.R

@@ -11,25 +11,27 @@
  
  insert_msg('Table 1: baseline cohort characteristic')
  
- paper_tables$baseline_cohort <- eda$baseline$summary_cohort %>% 
-    select(label, statistic) %>% 
-    map_dfc(stri_replace, regex = '\\ncomplete.*', replacement = '') %>% 
+ paper_tables$baseline_cohort <- ed_analysis$baseline %>% 
+    mutate(variable = translate_var(variable)) %>% 
+    map_dfc(stri_replace, regex = '\\nComplete.*', replacement = '') %>% 
     set_names(c('Variable', paste('Statistic, n =', nrow(radio$clear))))
  
 # Table 2: CT abnormalities at the visits, whole cohort ----
  
  insert_msg('Table 2: CT abnormalities, whole cohort, consecutive visits')
  
- paper_tables$ct_cohort <- eda$ct$summary_cohort %>% 
-    select(label, 
-           starts_with('fup'), 
-           significance) %>% 
+ paper_tables$ct_cohort <- ed_analysis$ct %>% 
+    filter(!variable %in% c('opacity', 'hiopacity', 'ards')) %>% 
+    mutate(variable = translate_var(variable)) %>% 
+    map_dfc(stri_replace, regex = 'no.*\\nyes:\\s{1}', replacement = '') %>% 
+    select(-test) %>% 
     set_names(c('Variable', 
                 '2-month FUP', 
                 '3-month FUP', 
                 '6-month FUP', 
                 '1-year FUP', 
-                'Significance'))
+                'Significance', 
+                'Effect size'))
  
 # Table S1: Study variables -----
  
@@ -52,74 +54,33 @@
  
  insert_msg('Table S2: baseline characteristic, severity groups')
  
- suppl_tables$baseline_severity <- eda$baseline$summary_severity %>% 
-    select(label, 
-           mild, 
-           moderate, 
-           severe, 
-           critical, 
-           significance) %>% 
-    map_dfc(stri_replace, regex = '\\ncomplete.*', replacement = '') %>% 
+ suppl_tables$baseline_severity <- ed_analysis$severity %>% 
+    mutate(variable = translate_var(variable)) %>% 
     set_names(c('Variable', 
-                map2_chr(c('Mild', 'Moderate', 'Severe', 'Critical'), 
-                         map_chr(eda$baseline$analyses_severity$age$stat_tables, ~.$n_complete), 
-                         ~paste0(.x, ', n = ', .y)), 
-                'Significance'))
+                'Mild', 
[…]

### 2022-02-08 render_paper.R

@@ -4,13 +4,19 @@
   
 # tools ----
   
-  c('./tools/sys_tools.R', 
-    './tools/project_tools.R', 
-    './tools/plotting_tools.R') %>% 
-    walk(source)
-  
   library(writexl)
+  library(figur) ## available from https://github.com/PiotrTymoszuk/figur
+  library(soucer) ## available from https://github.com/PiotrTymoszuk/soucer
+  library(cowplot)
+  library(rmarkdown)
+  library(knitr)
+  library(flextable)
+  library(bookdown)
   
+  c('./tools/project_tools.R', 
+    './tools/project_globals.R') %>% 
+    source_all(message = TRUE, crash = TRUE)
+
 # paper scripts -----
   
   insert_msg('Sourcing the paper scripts')

### 2022-02-08 tools/project_globals.R

@@ -46,7 +46,7 @@
                              'severe' = 'coral3', 
                              'critical' = 'firebrick3')  
   
-  globals$subset_labels <- c(cohort = 'Cohort', 
+  globals$subset_labels <- c(cohort = 'All participants', 
                              mild = 'Mild COVID-19', 
                              moderate = 'Moderate COVID-19', 
                              severe = 'Severe COVID-19', 

### 2022-02-08 tools/project_tools.R

@@ -238,264 +238,91 @@
     
   }
   
-# table formatting ------
-  
-  add_stat_info <- function(data_frame, signif_var = c('p_value', 'p_adj')) {
-    
-    ## adds information on tests and adjusts p values
-    
-    signif_var <- match.arg(signif_var[1], c('p_value', 'p_adj'))
-    
-    data_frame %>% 
-      mutate(p_adj = p.adjust(p_value, 'BH'), 
-             significance = format_p(.data[[signif_var]]), 
-             label = translate_var(variable), 
-             test = ifelse(class == 'factor', 'Chi squared', 'one-way ANOVA')) %>% 
-      map_dfc(function(x) if(is.character(x)) stri_replace(x, regex = '^.*\\nyes:\\s{1}', replacement = '') else x)
-    
-  }
-  
 # kinetic modeling and plotting -----
   
-  compare_effects <- function(data, 
-                              response = 'ctss', 
-                              indep_var = 'visit', 
-                              group_var = 'ID',
-                              invert = FALSE, 
-                              parallel = FALSE, 
-                              ci.type = 'norm') {
+  post_eff_size <- function(data) {
     
-    ## pairwise effect size
+    ## perforems post hoc testing with Wilcoxon test
+    ## and r effect size. Consecutive visits are compared
     
-    start_time <- Sys.time()
-    on.exit(message(paste('Elapsed:', Sys.time() - start_time)))
+    data <- data %>% 
+      dlply(.(visit))
     
-    pairs <- combn(unique(data[[indep_var]]), m = 2, simplify = FALSE)
+    test_lst <- list(fup1_fup2 = data[c('fup1', 'fup2')], 
+                     fup2_fup3 = data[c('fup2', 'fup3')], 
+                     fup3_fup4 = data[c('fup3', 'fup4')])
     
-    group_tbl <- pairs %>% 
-      map_dfr(~tibble(.y = response, 
-                      group1 = .x[1], 
-                      group2 = .x[2]))
+    p_values <- test_lst %>% 
+      map(~compare_variables(!!!.x, 
+                             variables = c('ctss', 'perc_opac'), 
+                             what = 'eff_size', 
+                             types = 'paired_wilcoxon_r', 
+                             ci = FALSE, 
+                             pub_styled = TRUE, 
+                             simplify_p = TRUE)) %>% 
+      map2_dfr(., names(.), ~mutate(.x, comparison = .y)) %>% 
[…]

### 2022-02-10 analysis scripts/risk_modeling_gam.R
Comaptibility with the newest versions of packages

@@ -190,7 +190,7 @@
   
   ## GAM models
   
-  risk_gam$comparison$gam_models <- list(caretx_model = risk_gam$cv$caret_models, 
+  risk_gam$comparison$gam_models <- list(x = risk_gam$cv$caret_models, 
                                          plot_title = list(c('GGO, GAM, training', 
                                                              'GGO, GAM, CV'), 
                                                            c('Reticulations, GAM, training', 
@@ -207,7 +207,7 @@
   
   ## Cutoff-models
   
-  risk_gam$comparison$cutoff_models <- list(caretx_model = risk$cv$caret_models, 
+  risk_gam$comparison$cutoff_models <- list(x = risk$cv$caret_models, 
                                             plot_title = list(c('GGO, age cutoff, training', 
                                                                 'GGO, age cutoff, CV'), 
                                                               c('Reticulations, age cutoff, training', 

### 2022-02-10 analysis scripts/score_modeling.R

@@ -59,7 +59,7 @@
   
   score$univariate$distr_plots <- map2(score$univariate$models, 
                                        c(10, 4), 
-                                       function(models, bin_no) list(lm_analysis_object = models, 
+                                       function(models, bin_no) list(x = models, 
                                                                      facet_formula = score$facet_forms) %>% 
                                          pmap(plot, 
                                               type = 'distribution', 
